---
title: "Downloading Census data"
author: "Sandrine Njoo & Dave Campbell"
date: "20/10/2021"
output: html_document
---
# Goal
Download 2016 census data about population counts in locations.  Then convert this document into a script to loop over provinces and extract dissemination area information from the census.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, warning = FALSE, message = FALSE}
#using R:
library(plyr)       # recoding factor levels using 'mapvalues'  
library(tidyverse)  # data manipulation and piping
library(data.table) # renaming columns in bulk using 'setnames'
library(sf)         # to handle census subdivision polygons

# set up data location:
location = "C:/Users/wher/OneDrive - Bank of Canada/MyDocs/git"

version
```

```{r, warning = FALSE, message = FALSE}
source('apikey.R')

```

## Set Up Province Code Mapping

Following the codings from [StatCan](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/tab/t1_8-eng.cfm).  This could be pulled directly from the html page, but it's a one time need and a small table so it is pasted below. 


```{r province_codings}

province_codings = matrix(c("Newfoundland and Labrador",	"NL",	"10",	"Atlantic",
      "Prince Edward Island",	"PE",	"11",	"Atlantic",
      "Nova Scotia",	"NS",	"12",	"Atlantic",
      "New Brunswick",	"NB",	"13",	"Atlantic",
      "Quebec",	"QC",	"24",	"Quebec",
      "Ontario",	"ON",	"35",	"Ontario",
      "Manitoba",	"MB",	"46",	"Prairies",
      "Saskatchewan",	"SK",	"47",	"Prairies",
      "Alberta",	"AB",	"48",	"Prairies",
      "British Columbia",	"BC",	"59",	"British Columbia",
      "Yukon",		"YT",	"60",	"Territories",
      "Northwest Territories",	"NT",	"61",	"Territories",
      "Nunavut",	"NU",	"62", "Territories"),
      ncol=4,byrow=TRUE, 
      dimnames = list(NULL, c("province", "abv", "code","region"))) %>%
  as_tibble(province_codings)


```


# Pull the 2016 census data for a province

```{r, cache = TRUE,  message = FALSE, warning=FALSE, quiet = TRUE}

#all the PEI regions by calling the province:
Province = "PE"




prov_code = province_codings %>% filter(abv == Province) %>% select(code)
regions=list(PR=prov_code)

#vectors that are relevant to industry
vectors_ind = c("v_CA16_5702","v_CA16_5705","v_CA16_5708","v_CA16_5711","v_CA16_5714","v_CA16_5717","v_CA16_5720","v_CA16_5723","v_CA16_5726","v_CA16_5729","v_CA16_5732","v_CA16_5735","v_CA16_5738","v_CA16_5741","v_CA16_5744","v_CA16_5747","v_CA16_5750","v_CA16_5753","v_CA16_5756","v_CA16_5759")

# pull from the census.  Use geo_format='sf' to also obtain the census boundary polygons and level = 'CSD' to extract the regions
census_data_ind <- get_census(dataset='CA16', regions=regions, vectors=vectors_ind, labels="detailed", geo_format='sf', level='CSD')

```



```{r, cache = TRUE, eval=FALSE,  message = FALSE, warning=FALSE}

#this is for all the possible variables in the 25% census data sample

#all the vectors for the 25% census
vectors_all=c("v_CA16_4951","v_CA16_4954","v_CA16_4960","v_CA16_4966","v_CA16_4972","v_CA16_4978","v_CA16_6751","v_CA16_6754","v_CA16_4984","v_CA16_4987","v_CA16_4990","v_CA16_4993","v_CA16_4996","v_CA16_4999","v_CA16_5002","v_CA16_5005","v_CA16_2601","v_CA16_2604","v_CA16_2613","v_CA16_2616","v_CA16_3390","v_CA16_3405","v_CA16_3438","v_CA16_3456","v_CA16_3636","v_CA16_3819","v_CA16_3831","v_CA16_3852","v_CA16_3879","v_CA16_3888","v_CA16_3954","v_CA16_3999","v_CA16_4840","v_CA16_4843","v_CA16_4849","v_CA16_4862","v_CA16_4870","v_CA16_4836","v_CA16_4856","v_CA16_4859","v_CA16_4873","v_CA16_4877","v_CA16_4886","v_CA16_4888","v_CA16_4890","v_CA16_4897","v_CA16_5051","v_CA16_5060","v_CA16_5063","v_CA16_5078","v_CA16_5096","v_CA16_5105","v_CA16_5108","v_CA16_5123","v_CA16_5141","v_CA16_5147","v_CA16_5153","v_CA16_5162","v_CA16_5189","v_CA16_5213","v_CA16_5225","v_CA16_5243","v_CA16_5258","v_CA16_5282","v_CA16_5291","v_CA16_5303","v_CA16_5321","v_CA16_5327","v_CA16_5333","v_CA16_5339","v_CA16_5348","v_CA16_5375","v_CA16_5399","v_CA16_5411","v_CA16_5429","v_CA16_5444","v_CA16_5468","v_CA16_5477","v_CA16_5489","v_CA16_5507","v_CA16_5513","v_CA16_5519","v_CA16_5522","v_CA16_5531","v_CA16_5555","v_CA16_5561","v_CA16_5564","v_CA16_5573","v_CA16_5597","v_CA16_5600","v_CA16_5621","v_CA16_5627","v_CA16_5639","v_CA16_5645","v_CA16_5654","v_CA16_5660","v_CA16_5693","v_CA16_5699","v_CA16_5762","v_CA16_5867","v_CA16_5870","v_CA16_6119","v_CA16_6659","v_CA16_5777","v_CA16_5792","v_CA16_5813","v_CA16_5831","v_CA16_6692","v_CA16_6698","v_CA16_6704","v_CA16_6707","v_CA16_6719","v_CA16_6725","v_CA16_6731","v_CA16_6734")

# pull from the census.  Use geo_format='sf' to also obtain the census boundary polygons
census_data_all <-  get_census(dataset='CA16', regions=regions, vectors=vectors_all, labels="detailed", geo_format='sf', level='CSD')

```

```{r renamingvariables, cache = TRUE}
# rename variables to get rid of non-alphanumeric characters
names(census_data_ind) <- str_replace_all(names(census_data_ind), pattern="v_CA\\d+_\\d+:\\s", replacement = "X")
census_data_ind = census_data_ind %>% rename_all(make.names)
census_data_ind %>% glimpse
```


Adjust some industries so they match GDP and THW datasets
```{r}
census_data_ind = census_data_ind %>% 
  mutate(X11.Agriculture.forestry.fishing.hunting.21.Mining.quarrying.and.oil.and.gas.extraction = X11.Agriculture..forestry..fishing.and.hunting + X21.Mining..quarrying..and.oil.and.gas.extraction) %>%
  mutate(X41.Wholesale.trade.44.45.Retail.trade                                                  = X41.Wholesale.trade+ X44.45.Retail.trade) %>%
  mutate(X52.Finance.and.insurance.53.Real.estate.and.rental.and.leasing                         = X52.Finance.and.insurance + X53.Real.estate.and.rental.and.leasing) %>%
  mutate(X55.Business.building.and.other.support.services.56                                     = X55.Management.of.companies.and.enterprises + X56.Administrative.and.support..waste.management.and.remediation.services) %>%
  mutate(X51.Information.culture.and.recreation.71                                               = X51.Information.and.cultural.industries + X71.Arts..entertainment.and.recreation)


```


There are a few options for making plots.  See also [here](https://mountainmath.github.io/cancensus/articles/Making_maps_with_cancensus.html) for interactive maps via **leaflet**.
```{R makeplots}
#base R plots:
plot(census_data_ind["X41.Wholesale.trade.44.45.Retail.trade"], main = "LFS counts of individuals: PEI wholesale and retail trade")


# ggplot maps:
census_data_ind %>% ggplot() +
  geom_sf(aes(fill = X41.Wholesale.trade.44.45.Retail.trade))

```



```{r savingoutput}

# construct a file name that is both descriptive and based on the name of the file used to construct and manipulate the data.
census_data_ind  %>% write_csv(file = paste0(location,'1.a-Census2016-data/1.a_census_data_',Province,'.csv'))
census_data_ind  %>% write_rds(file = paste0(location,'1.a-Census2016-data/1.a_census_data_',Province,'.rds'))
province_codings %>% write_csv(file = paste0(location,'1.a_province_codings.csv'))

```


# As a script and download provincial Dissemination Areas

Convert this markdown into a script by running the below line.  Loop over provinces and select **Dissemination Areas** rather than **Census Subdivisions**.  For clarity this is pasted below.
```{r, eval = FALSE}
knitr::purl("1.a-Census2016-download.Rmd")
```

### Loop over provinces and extract Dissemination Areas.

```{r, message = FALSE, eval = TRUE}
#Loop over all provinces:
for(Province in province_codings$abv){
  
  prov_code = province_codings %>% filter(abv == Province) %>% select(code)
  regions=list(PR=prov_code)
  
  # pull from the census.  Use geo_format='sf' to also obtain the census boundary polygons and level = 'CSD' to extract the regions
  census_data_ind <- get_census(dataset='CA16', regions=regions, vectors=vectors_ind, labels="detailed", geo_format='sf', level='DA')
  
  # rename variables to get rid of non-alphanumeric characters
  names(census_data_ind) <- str_replace_all(names(census_data_ind), pattern="v_CA\\d+_\\d+:\\s", replacement = "X")
  census_data_ind = census_data_ind %>% rename_all(make.names)
  
  census_data_ind = census_data_ind %>% 
    mutate(X11.Agriculture.forestry.fishing.hunting.21.Mining.quarrying.and.oil.and.gas.extraction = X11.Agriculture..forestry..fishing.and.hunting + X21.Mining..quarrying..and.oil.and.gas.extraction) %>%
    mutate(X41.Wholesale.trade.44.45.Retail.trade                                                  = X41.Wholesale.trade+ X44.45.Retail.trade) %>%
    mutate(X52.Finance.and.insurance.53.Real.estate.and.rental.and.leasing                         = X52.Finance.and.insurance + X53.Real.estate.and.rental.and.leasing) %>%
    mutate(X55.Business.building.and.other.support.services.56                                     = X55.Management.of.companies.and.enterprises + X56.Administrative.and.support..waste.management.and.remediation.services) %>%
    mutate(X51.Information.culture.and.recreation.71                                               = X51.Information.and.cultural.industries + X71.Arts..entertainment.and.recreation) %>%
    mutate(PR_UID = province_codings %>% filter(abv == Province)%>% pull(code)) # add in province code so that DA and CSD match.
  
  # construct a file name that is both descriptive and based on the name of the file used to construct and manipulate the data.
  census_data_ind  %>% write_csv(file = paste0(location,'1.a-Census2016-data/1.a_census_data_',Province,'_DA.csv'))
  census_data_ind  %>% write_rds(file = paste0(location,'1.a-Census2016-data/1.a_census_data_',Province,'_DA.rds'))
  
}

```
