---
title: "Obtain and clean up the LFS data"
author: "Dave Campbell"
date: "08/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



# Goal:
Acquire the Labour Force Survey microdata.  Format and rename variables.  Loop over all months of 2021.  Previous years seem to be available but must be ordered  StatCan.

### Note: Not Used Elsewhere



## Software Version


```{r Rversion}
#using R:
version
```

## Import libraries


```{r library, warning = FALSE, message = FALSE}
#using R:
library(tidyverse)  # data manipulation and piping
library(data.table) # renaming columns in bulk using 'setnames'
library(plyr)       # recoding factor levels using 'mapvalues'  
library(glue)       # used for dynamic variable recoding

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```





## Data Documentation and Direct Data Download.


Here we use the monthly *Labour Force Survey Public Use Microfile* from [StatCan](https://www150.statcan.gc.ca/n1/pub/71m0001x/71m0001x2021001-eng.htm). This file directly downloads a zip file from Statistics Canada containing both the metadata and csv datafile.


Downloading LFS directly from StatCan.  



```{r, cache = TRUE,  message = FALSE, warning=FALSE}

# make a temp file
lfs_data = NULL

year  = "2021"
for(month in c("01","02","03","04","05","06","07","08","09","10","11","12")){
  temp <- tempfile()
  # note the change in formatting of the url is handled here:
  if(month %in% c("01","02","03","04")){
    month2use = paste0(month, "-CSV-eng")
  }else{
    month2use = paste0(month, "-CSV")
  }
  #construct the url 
  url = paste0("https://www150.statcan.gc.ca/n1/en/pub/71m0001x/2021001/",year,"-",month2use,".zip")
  download.file(url,temp) # download the file from the url into the temp file
  (file_list <- as.character(unzip(temp, list = TRUE)$Name)) #list = TRUE <-- list the files but do not exract them
  # extract the metadata
  lfs_meta_data <- read_csv(unz(temp, file_list[1]),skip = 1, col_names = FALSE) 
  # Here I skip the first line since it contains '\' which is read as an invalid character. 
  # Since I've skipped the column names in the csv I tell it not to make column names
  tmp_data      <- read_csv(unz(temp, file_list[3])) %>%
    rename_all(make.names)%>% # rebuild names so that column names no longer include spaces and non-alphanumeric characters
    mutate(year = year, month = month) 
  unlink(temp) # delete the temp file
  lfs_data = rbind(lfs_data,tmp_data)
}

# # showing the data files
# # meta data file:
# lfs_meta_data %>% head
# # datafile is painfully unreadable:
# lfs_data %>% glimpse()


```

## Making more descriptive variable names from metadata.
Recode CMA and Province names into R friendly values.  This improves readibility, prevents errors in filtering, and simplifies merging with other data sets later on.

**Assuming** that LFS metadata didn't change substantially...

```{r renamingvariables, cache = TRUE}

# recode province and CMA by extracting coding from the metadata by first finding where in the metadata file the codings are kept:
province_index = which(str_detect(lfs_meta_data$X5, pattern = "prov"))
cma_index      = which(str_detect(lfs_meta_data$X5, pattern = "cma"))

province_codings = tibble(old_names = as.numeric(lfs_meta_data$X12[province_index+c(1:10)]),
                         new_names = lfs_meta_data$X13[province_index+c(1:10)])

cma_codings      = tibble(old_names = as.numeric(lfs_meta_data$X12[cma_index+c(1:10)]),
                         new_names = lfs_meta_data$X13[cma_index+c(1:10)])

# We can also clean up the poor encoding of accent letters or deal with spaces

cma_codings = cma_codings %>% mutate(new_names = iconv(cma_codings$new_names, to='UTF-8'))%>% # get rid of poor encoding
                              mutate(new_names = str_replace_all(new_names, pattern = "\xe9"                  , replacement = "e"))%>% # swap "é" with "e" in case this causes problems for my English keyboard later.
                              mutate(new_names = str_replace_all(new_names, pattern = "\\s|[[:punct:]]|(\x96)", replacement = "_")) # spaces, punctuation or badly encoded slashes.
province_codings = province_codings %>% 
                             mutate(new_names = str_replace_all(new_names, pattern = "\\s",                     replacement = "_"))

# resulting province and cma names:
province_codings
cma_codings


# Recoding provinces and cmas:
lfs_data = lfs_data %>%
  mutate(PROV = plyr::mapvalues(PROV, from = province_codings$old_names, to = province_codings$new_names))%>%
  mutate(CMA  = plyr::mapvalues(CMA,  from = cma_codings$old_names,      to = cma_codings$new_names))

# build a mapping from old names to the more descriptive names:
new_variable_name_recodings = lfs_meta_data %>% 
  mutate(old_name = toupper(X5), new_name = str_replace_all(toupper(X6), pattern = "\\s+|[:punct:]+",replacement = "_"))%>%
  mutate(new_name =                         str_replace_all(new_name,    pattern = "_{2,}",          replacement = "_"))%>%
  select(old_name, new_name)%>% 
  drop_na()
# This makes most variable names easier to read, but 
# a few new_variable names are not unique.  
# Option 1: just alter the duplicate variable names
# Option 2: combine all the old and new names.  <-- This is probably more robust and consistent although it makes for longer names
#          # A user could then search for the original or altered variable names easily


# # Option 1: just alter the duplicate variable names
# 
# duplicatenames = names(which(table(new_variable_name_recodings$new_name)>1))
# 
# 
# for(duplicate in duplicatenames){
# index = which(new_variable_name_recodings$new_name ==duplicate)
#   new_variable_name_recodings$new_name[index]  = 
#     paste(new_variable_name_recodings$old_name[index], 
#           new_variable_name_recodings$new_name[index], sep=".")
#     
# }

# Option 2: combine all the old and new names.  <-- This is probably more future proof 
#
new_variable_name_recodings = new_variable_name_recodings %>%
                        mutate(new_name = paste(old_name, new_name, sep="."))




# change the variable names to (longer) more descriptive names
setnames( lfs_data,
old = new_variable_name_recodings$old_name, 
new = new_variable_name_recodings$new_name, skip_absent = TRUE)




```


## General Variable Recoding From MetaData

Recode all the variables dynamically by extracting the recodings from the metadata.  This also re-scales numeric variables which have one or two implied decimal places.


```{r recode_variables_dynamically, warning = FALSE, message = FALSE}
# set up a new data frame:
lfs_data_recode = lfs_data


#Loop over variable names
# recode factors using the metadata
# rescale cases where the variable is numeric and has 1 or 2 implied decimals

# cycle through almost all of the variables except these ones:
colnames2use = setdiff(colnames(lfs_data_recode), 
                        c("REC_NUM.ORDER_OF_RECORD_IN_FILE","SURVYEAR.SURVEY_YEAR","SURVMNTH.SURVEY_MONTH","PROV.PROVINCE","CMA.NINE_LARGEST_CMAS", "FINALWT.STANDARD_FINAL_WEIGHT", "year","month"))

for(variable_from_lfs_data in colnames2use){


  ## find the relevant rows of the metadata relating to this variable
  metadata_all_vars_index = which(!is.na(lfs_meta_data$X5)) # starting point for all variables
  metadata_index_start = which(str_detect(
        tolower(lfs_meta_data$X5), 
        pattern = tolower(unlist(strsplit(variable_from_lfs_data,split="\\."))[1])
    ))# starting point for our variable
  metadata_index_end = metadata_all_vars_index[metadata_all_vars_index>metadata_index_start][1] #starting point for the subsequent variable

  # extract the lfs_data coded variable and its meaning
  data_codings = tibble(old_names = lfs_meta_data$X12[metadata_index_start:metadata_index_end],
                        new_names = lfs_meta_data$X13[metadata_index_start:metadata_index_end])%>%
                 drop_na()
  # We can also clean up the poor encoding of accent letters or deal with spaces

  data_codings = data_codings %>% mutate(new_names = iconv(data_codings$new_names, to='UTF-8'))%>% #  poor encoding
                              mutate(new_names = str_replace_all(new_names, 
                                                                 pattern = "\xe9",                   replacement = "e"))%>% # swap "é" with "e" 
                              mutate(new_names = str_replace_all(new_names, 
                                                                 pattern = "\\s|[[:punct:]]|(\x96)", replacement = "_")) # spaces, punctuation or badly encoded slashes.

  
    # recode the categorical variables, some variables are numeric.  These won't be recoded here, see "else"
    if(nrow(data_codings)>1 & 
        data_codings[1,1]!="1-990" &  # hours per week worked is not recoded in this way 
        data_codings[1,1]!="0-990" &  # hours away from work is not recoded in this way
        data_codings[1,1]!="1-999999" # money earned is not recoded in this way
      ){
      
    # Recoding variables:
    # Using glue we can dynamically create the line of code to recode the variable.
    # eval(parse(text=... treats the 'glue' code as text that is to be evaluated.
    # note that values within {} are evaluated so that they are placed into the code to excecute
    # The way this is coded I could have just used 'paste' to do this.  

    eval(parse(text= glue("lfs_data_recode = lfs_data_recode %>% mutate({variable_from_lfs_data} = mapvalues({variable_from_lfs_data}, from = data_codings$old_names, to = data_codings$new_names))")))
    }else{ # handling variables with one or two decimals:
     # handling 1 decimal place recodings:
      if(nrow(data_codings)>1 & any(str_detect(data_codings$new_names, pattern = "implied_decimal"))){ 
        # Recoding variables:
        # Using glue we can dynamically create the line of code to recode the variable.
        # eval(parse(text=... treats the 'glue' code as text that is to be evaluated.
        # note that values within {} are evaluated so that they are placed into the code to excecute
        # The way this is coded I could have just used 'paste' to do this.  

        eval(parse(text= glue("lfs_data_recode = lfs_data_recode %>% mutate({variable_from_lfs_data} = as.numeric({variable_from_lfs_data})/10)")))
        }else{
        # handling 2 decimal place recodings:
          if(nrow(data_codings)>1 & any(str_detect(data_codings$new_names, pattern = "two_decimals"))){ 
          # Recoding variables:
          # Using glue we can dynamically create the line of code to recode the variable.
          # eval(parse(text=... treats the 'glue' code as text that is to be evaluated.
          # note that values within {} are evaluated so that they are placed into the code to excecute
          # The way this is coded I could have just used 'paste' to do this.  
  
          eval(parse(text= glue("lfs_data_recode = lfs_data_recode %>% mutate({variable_from_lfs_data} = as.numeric({variable_from_lfs_data})/100)")))
          }
        }
    }
}


```

### Comparing files

```{r}
#original with renamed variables
lfs_data
#with renamed variables and recoded values
lfs_data_recode

```




### Saving output
Construct a descriptive name using the year = `r print(year)` as defined above.  Save both versions of the LFS with and without recoded variables.  

**Come back later and obtain more LFS data?**

```{r savingoutput}

# construct a file name that is both descriptive and based on the name of the file used to construct and manipulate the data.
write_csv(lfs_data, file = paste0(location,"1.b-lfs-data-from-",year,".csv"))
write_csv(lfs_data_recode, file = paste0(location,"1.b-lfs_data_recoded_vars-from-",year,".csv"))


```

### Saving Just PEI
Both versions of the dataset.
```{r pei, eval = TRUE}

# construct a file name that is both descriptive and based on the name of the file used to construct and manipulate the data.
province_selected = "Prince_Edward_Island"
lfs_data %>% filter(PROV.PROVINCE == province_selected) %>%
                    write_csv( file = paste0(location,"1.b-lfs-",province_selected,"-",year,".csv"))



lfs_data_recode %>% filter(PROV.PROVINCE == province_selected) %>%
                    write_csv( file = paste0(location,"1.b-lfs_recoded_vars-",province_selected,"-",year,".csv"))



```

