---
title: 'GDP and Total Hours Worked'
author: "Dave Campbell"
date: '2021-11-01'
summary: 'Cansim data extraction'
lastmod: '2022-09-06T16:34:12-04:00'
---

# Goals

Extract **Total Hours Worked** in each NAICS sector at \{province, monthly, NAICS sector\} and **GDP** at both {provincial, annual, NAICS sector} and \{national, monthly, NAICS sector\}.  Ultimately want to use total hours worked to split the GDP per sector into shares for each province.

# Limitations

- **Management.of.companies.and.enterprises..55.** is missing before 2007.  May need imputing by growing it back by total GDP and/or considering growth of other variables.

# Data Provider

Statistics Canada hosts a large set of socio-economic [datasets](https://www150.statcan.gc.ca/n1//en/type/data?MM=1#tables). 


# Basic libraries and R version

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(dplyr)  # for data manipulation

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```

```{R version}
version
```

  
## Pulling the raw tables from the cansim library

The **total hours wored** data table has 10 million rows and the **GDP** table has 1/4 milllion.  While we're at it, remake some column names to make them more R friendly and construct proper dates.


```{R, message = FALSE, cache = TRUE, warning = FALSE, eval = FALSE}
#Actual hours worked by industry, monthly, unadjusted for seasonality
THW = get_cansim("14-10-0036-01") %>% rename_all(make.names) 

#Actual hours worked by industry, monthly, adjusted for seasonality, national
THWSA = get_cansim("14-10-0289-01") %>% rename_all(make.names) 

#Employment by industry, monthly, unadjusted for seasonality
EMP = get_cansim("14-10-0355-01") %>% rename_all(make.names) 

#Gross domestic product (GDP) at basic prices, by industry, monthly (x 1,000,000)
GDP       = get_cansim("36-10-0434-01") %>% rename_all(make.names) 

#Gross domestic product (GDP) at basic prices, by industry, provinces and territories (x 1,000,000) annual
GDPAnnual = get_cansim("36-10-0402-01") %>% rename_all(make.names) 

#Gross domestic product (GDP) at factor cost in 1992 constant prices, by Standard Industrial Classification, 1980 (SIC), monthly (x 1,000,000)
GDPsic = get_cansim("36-10-0377-01") %>% rename_all(make.names)

```


### Basic date cleanup

```{R, cache = TRUE, eval = FALSE}
THW = THW %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month. <-- already included
  separate(REF_DATE, into = c("year","month"),sep = "-")

THWSA = THWSA %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month. <-- already included
  separate(REF_DATE, into = c("year","month"),sep = "-")

EMP = EMP %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month. <-- already included
  separate(REF_DATE, into = c("year","month"),sep = "-")

GDP = GDP %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month.<-- already included
  separate(REF_DATE, into = c("year","month"),sep = "-")

#GDPAnnual = GDPAnnual %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month.<-- already included
 # mutate(year= as.numeric(year)) - this seems to not do anything? 

GDPsic = GDPsic %>% 
  # mutate(date = parse_date_time(REF_DATE, orders = "Y-m")) %>%# 4 digit year with 2 digit month.<-- already included
  separate(REF_DATE, into = c("year","month"),sep = "-")

save.image("Total_hours_and_GDP.Rdata")

```


```{r loaddata, echo = FALSE}
# loading from here to avoid re-downloading
load("Total_hours_and_GDP.Rdata")
```

### Widen the table so that each row is a single date

Starting with Total Hours Worked, note the filtering to **Both sexes** and **Total actual hours (main job)** so that rows have unique values for a given date and location.

```{r, widendata,  cache = TRUE}


THW_wide = THW%>%filter(Sex == "Both sexes") %>% 
  filter(Class.of.worker == "Total employed")%>%
  filter(Actual.hours.worked == "Total actual hours (main job)")%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Goods.producing.sector" )%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Services.producing.sector" )%>%
  #filter(GEO != "Canada")%>% # since the Canada value will be a type of weighted average
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS.,
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the province names


THWSA_wide = THWSA%>% 
  filter(Statistics == "Estimate")%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Goods.producing.sector" )%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Services.producing.sector" )%>%
  #filter(GEO != "Canada")%>% # since the Canada value will be a type of weighted average
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS.,
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the province names


# Employment - need seasonally adjusted and non-seasonally adjusted

EMPSA_wide = EMP%>% 
  filter(Statistics == "Estimate")%>%
  filter(Data.type == "Seasonally adjusted")%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Goods.producing.sector" )%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Services.producing.sector" )%>%
  #filter(GEO != "Canada")%>% # since the Canada value will be a type of weighted average
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS.,
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the province names

##

EMPNSA_wide = EMP%>%
  filter(Statistics == "Estimate")%>%
  filter(Data.type == "Unadjusted")%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Goods.producing.sector" )%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Services.producing.sector" )%>%
  #filter(GEO != "Canada")%>% # since the Canada value will be a type of weighted average
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS.,
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the province names



```


## 2012 constant or chained dollars
Note the decisions that need to be made for GDP include filtering on some variables.  Consider these price options:
```{r exploreGDP, cache=TRUE}
GDP %>% 
  filter(GEO =="Canada")%>%
  filter(Seasonal.adjustment=="Seasonally adjusted at annual rates")%>%
  select(Prices)%>% table %>% names %>% sort

```

But the options available differ when considering the annual GDP per sector, per province:
```{r exploreGDPAnnual, cache=TRUE}
GDPAnnual %>% 
  filter(GEO =="Canada")%>%
  select(Value)%>% table %>% names %>% sort

```

To be consistent it will force us to select **Chained (2012) dollars**.  
```{r}
  dollars2use = "Chained (2012) dollars"
# dollars2use = "2012 constant prices"
```

## NAICS
There are a lot of different values in the NAICS in the GDP datasets, fortunately they aling between the two GDP datasets.  Here are 15 of them:
```{r exploreGDP2, cache=TRUE}

GDP %>% 
  filter(GEO =="Canada")%>%
  filter(Seasonal.adjustment=="Seasonally adjusted at annual rates")%>%
  filter(Prices==dollars2use)%>%
  select(  North.American.Industry.Classification.System..NAICS.)%>% 
  table %>% names %>% sort %>% head(15)

#GDPretro %>% 
#  filter(GEO =="Canada")%>%
#  filter(Seasonal.adjustment=="Seasonally adjusted")%>%
#  select(Standard.Industrial.Classification..1980..SIC.)%>% 
#  table %>% names %>% sort %>% head(15)

# 
# GDPAnnual %>% 
#   filter(GEO =="Canada")%>%
#   filter(Value==dollars2use)%>%
#   select(North.American.Industry.Classification.System..NAICS.)%>% 
#   table %>% names %>% sort %>% head(15)

```


Reducing down to a smaller set of of values from the NAICS hierarchy.

```{r andrevalues, cache = TRUE}

NAICS_higher_levels = c(
"All industries [T001]",
"Agriculture, forestry, fishing and hunting [11]",
"Mining, quarrying, and oil and gas extraction [21]",	
"Utilities [22]",
"Construction [23]",
"Manufacturing [31-33]",
"Wholesale trade [41]",
"Retail trade [44-45]",
"Transportation and warehousing [48-49]",
"Information and cultural industries [51]",
"Finance and insurance [52]",
"Real estate and rental and leasing [53]",
"Professional, scientific and technical services [54]",
"Management of companies and enterprises [55]",
"Administrative and support, waste management and remediation services [56]",
"Educational services [61]",
"Health care and social assistance [62]",
"Arts, entertainment and recreation [71]",
"Accommodation and food services [72]",
"Other services (except public administration) [81]",
"Public administration [91]")

GDP_wide = 
  GDP %>% 
  filter(Seasonal.adjustment=="Seasonally adjusted at annual rates")%>%
  filter(Prices==dollars2use)%>%
  filter(North.American.Industry.Classification.System..NAICS. %in%NAICS_higher_levels)%>%
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the NAICS names

GDPAnnual_wide = 
  GDPAnnual %>% 
  filter(Value==dollars2use)%>%
  filter(North.American.Industry.Classification.System..NAICS. %in%NAICS_higher_levels)%>%
  select(GEO,Date, 
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the NAICS names



```

Selecting SIC codes that can be matched with NAICS

```{r andrevalues, cache = TRUE}
SIC_NAICS_common = c(
  "All industries",
  "Agricultural and related services industries",
  "Fishing and trapping industries",
  "Logging and forestry industries",
  "Mining (including milling), quarrying and oil well industries",
  "Manufacturing industries",
  "Construction industries",
  "Transportation and storage industries",
  "Communication industries",
  "Other utility industries",
  "Wholesale trade industries",
  "Retail trade industries",
  "Finance, insurance and real estate industries",
  "Business services industries",
  "Government service industries",
  "Educational service industries",
  "Health and social service industries",
  "Accommodation, food and beverage service industries",
  "Other service industries",
  "Amusement and recreational service industries",
  "Postal and courier service industries")
  
GDPsic_wide = 
  GDPsic %>% 
  filter(Seasonal.adjustment=="Seasonally adjusted")%>%
  filter(Standard.Industrial.Classification..1980..SIC. %in% SIC_NAICS_common)%>%
  select(GEO,Date,year, month,
               Standard.Industrial.Classification..1980..SIC.,
               VALUE)%>%
  pivot_wider(names_from= Standard.Industrial.Classification..1980..SIC., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the SIC names

```



## making THW and GDP comparable

Some NAICS categories do not completely overlap across the THW and GDP datasets.  Combine categories accordingly.

```{r}

GDP_wide = GDP_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21   =  Agriculture..forestry..fishing.and.hunting..11. +Mining..quarrying..and.oil.and.gas.extraction..21.)%>%
  mutate(Wholesale.and.retail.trade..41..44.45.                     = Wholesale.trade..41.+Retail.trade..44.45.)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. = Finance.and.insurance..52.+Real.estate.and.rental.and.leasing..53.)%>%
  mutate(Information..culture.and.recreation..51..71.                = Arts..entertainment.and.recreation..71.+Information.and.cultural.industries..51.)


GDPAnnual_wide = GDPAnnual_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21   =  Agriculture..forestry..fishing.and.hunting..11. +Mining..quarrying..and.oil.and.gas.extraction..21.)%>%
  mutate(Wholesale.and.retail.trade..41..44.45.                     = Wholesale.trade..41.+Retail.trade..44.45.)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. = Finance.and.insurance..52.+Real.estate.and.rental.and.leasing..53.)%>%
  mutate(Professional..scientific.and.technical.services..54.       = Management.of.companies.and.enterprises..55.+Administrative.and.support..waste.management.and.remediation.services..56.+Professional..scientific.and.technical.services..54.)%>%
  mutate(Information..culture.and.recreation..51..71.                = Arts..entertainment.and.recreation..71.+Information.and.cultural.industries..51.)


GDPsic_wide = GDPsic_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21 = Agricultural.and.related.services.industries + Fishing.and.trapping.industries + Logging.and.forestry.industries + Mining..including.milling...quarrying.and.oil.well.industries)%>%
  mutate(Wholesale.and.retail.trade..41..44.45. = Wholesale.trade.industries + Retail.trade.industries)%>%
  mutate(Transportation.and.warehousing..48.49. = Transportation.and.storage.industries + Postal.and.courier.service.industries)%>%
  mutate(Information..culture.and.recreation..51..71. = Communication.industries - Postal.and.courier.service.industries + Amusement.and.recreational.service.industries) %>%
mutate(Other.service.industries = Other.service.industries - Amusement.and.recreational.service.industries)


```


### Sanity check for entire sample

```{r}

GDP.variables = c("Utilities..22.",
      "Construction..23.",                                      
      "Manufacturing..31.33.",  
      "Transportation.and.warehousing..48.49.", 
      "Educational.services..61.",                      
      "Health.care.and.social.assistance..62.",
      "Accommodation.and.food.services..72.",
      "Other.services..except.public.administration...81.",
      "Public.administration..91.",
      "Agriculture..forestry..fishing.and.hunting..11.mining.21",
      "Wholesale.and.retail.trade..41..44.45.",
      "Finance..insurance..real.estate..rental.and.leasing..52.53.",
      "Professional..scientific.and.technical.services..54.", 
      "Information..culture.and.recreation..51..71.")

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDP_post1996 = GDP_wide %>%  mutate(All.industries.ours = ", paste({GDP.variables},collapse='+'),")"))
           )
     )


#plot to make sure that our combination = the stated combination
GDP_post1996%>% ggplot()+
  geom_point(aes(x=All.industries.ours, y=All.industries..T001., colour = year))



eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDPAnnual_post1996 = GDPAnnual_wide %>%  mutate(All.industries.ours = ", paste({GDP.variables},collapse='+'),")"))
           )
     )

#now for GDPsic - replace these with SIC codes

GDPSIC.variables = c("Manufacturing.industries", "Construction.industries", "Other.utility.industries", "Finance..insurance.and.real.estate.industries", "Business.services.industries", "Government.service.industries", "Educational.service.industries", "Health.and.social.service.industries", "Accommodation..food.and.beverage.service.industries", "Other.service.industries", "Agriculture..forestry..fishing.and.hunting..11.mining.21", "Wholesale.and.retail.trade..41..44.45.", "Transportation.and.warehousing..48.49.", "Information..culture.and.recreation..51..71.")

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDP_pre1997 = GDPsic_wide %>%  mutate(All.industries.ours = ", paste({GDPSIC.variables},collapse='+'),")"))
           )
     )

#plot to make sure that our combination = the stated combination
GDP_pre1997%>% ggplot()+
  geom_point(aes(x=All.industries.ours, y=All.industries, colour = year))

```

##need to impute values for pre-2007 management - NAICS possible solutions include (1) grow it backwards at the rate of GDP growth (2) regress management 2007- on GDP, use estimated relationship to impute (3) estimate relationship between management and all industries 


```{r}

#(1)
#value of management in dec 2006 x 1 - growth rate of gdp in January 2007 

#create variable for GDP growth rate:
GDP_post1996 = GDP_post1996 %>% 
   arrange(Date) %>%
mutate(GDP_growth=All.industries..T001./lag(All.industries..T001.)-1) 

#(1) Looping backward from 2006-12-01:

start     <- as.Date("06-12-01",format="%y-%m-%d")
end   <- as.Date("97-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_post1996 = GDP_post1996 %>% mutate(Management.of.companies.and.enterprises..55. = case_when( (Date < as.Date("2007-01-01")) ~ lead(Management.of.companies.and.enterprises..55.)*(1-lead(GDP_growth)),(Date >= as.Date("2007-01-01")) ~ Management.of.companies.and.enterprises..55.))
  theDate <- theDate - 1}

GDP_post1996 = GDP_post1996 %>% 
 mutate(Professional..scientific.and.technical.services..54.       = Management.of.companies.and.enterprises..55.+Administrative.and.support..waste.management.and.remediation.services..56.+Professional..scientific.and.technical.services..54.)

 eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDP_post1996 = GDP_post1996 %>%  mutate(All.industries.ours = ", paste({GDP.variables},collapse='+'),")"))
           )
     )
 


#we now have a complete GDP_post1996 table with no NA values
 

 
```


### Check which variables were now left out from crafting our total
Compare these to the list in the previous chunk that was used to build **totals.hours.ours**.  Note that above, some of these _leftover_ variables were already used (eg. mining + agriculture).

```{r}
# for book keeping
setdiff(colnames(GDP_post1996),GDP.variables)
setdiff(colnames(GDP_pre1997),GDPSIC.variables)
```


### This time with Total Hours Worked and employment

```{r}


THW.variables = c(
  "Agriculture..111.112..1100..1151.1152.",
  "Forestry..fishing..mining..quarrying..oil.and.gas..21..113.114..1153..2100.",
  "Utilities..22.",
  "Construction..23.",
  "Manufacturing..31.33.",
  "Wholesale.and.retail.trade..41..44.45.",
  "Transportation.and.warehousing..48.49.",
  "Finance..insurance..real.estate..rental.and.leasing..52.53.",
  "Professional..scientific.and.technical.services..54.", 
  "Business..building.and.other.support.services..55.56.",
  "Educational.services..61.",
  "Health.care.and.social.assistance..62.",
  "Information..culture.and.recreation..51..71.",               
  "Accommodation.and.food.services..72.",
  "Other.services..except.public.administration...81.",             
  "Public.administration..91.")


eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("THW_post1986 = THW_wide %>% mutate(totals.hours.ours = ", paste({THW.variables},collapse='+'),")"))
           )
     )

#plot to make sure that our combination = the stated combination
THW_post1986 %>% ggplot()+
  geom_point(aes(x=totals.hours.ours, y=Total.employed..all.industries, colour = GEO))+
  geom_abline(slope=1,intercept=0)

THW_post1986 <- THW_post1986 %>% select(-contains(".sector"))

# Repeat for others:

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("THWSA_post1986 = THWSA_wide %>% mutate(totals.hours.ours = ", paste({THW.variables},collapse='+'),")"))
           )
     )


THWSA_post1986 <- THWSA_post1986 %>% select(-contains(".sector"))


#

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("EMPSA_post1986 = EMPSA_wide %>% mutate(totals.emp.ours = ", paste({THW.variables},collapse='+'),")"))
           )
     )

EMPSA_post1986 <- EMPSA_post1986 %>% select(-contains(".sector"))

#

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("EMPNSA_post1986 = EMPNSA_wide %>% mutate(totals.emp.ours = ", paste({THW.variables},collapse='+'),")"))
           )
     )

EMPNSA_post1986 <- EMPNSA_post1986 %>% select(-contains(".sector"))


```

```{r}
# for book keeping
setdiff(colnames(THW_post1986),THW.variables)
```
## We need to create a seasonal factor for employment 

```{r}


EMPSA1 <- subset (EMPSA_post1986, select = -c(GEO,Date,year,month))
EMPNSA1 <- subset (EMPNSA_post1986, select = -c(GEO,Date,year,month))
info_columns <- subset (EMPSA_post1986, select = c(GEO,Date,year,month))

EMPSF_post1986 <- cbind(info_columns, EMPSA1/EMPNSA1)

EMPSF_post1986 <- subset(EMPSF_post1986, Date > as.Date("1986-12-01"))



```

## We need a seasonal factor for THW for Canada

```{r}
# Create seasonal factor for Canada by hours

THWNSA_wide = THW%>%filter(Sex == "Both sexes") %>% 
  filter(Class.of.worker == "Total employed")%>%
  filter(Actual.hours.worked == "Total actual hours (main job)")%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Goods.producing.sector" )%>%
  filter(North.American.Industry.Classification.System..NAICS.!= "Services.producing.sector" )%>%
  filter(GEO == "Canada")%>% 
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS.,
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the province names

eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("THWNSA_post1986 = THWNSA_wide %>% mutate(totals.hours.ours = ", paste({THW.variables},collapse='+'),")"))
           )
     )

THWNSA_post1986 <- THWNSA_post1986 %>% select(-contains(".sector"))


THWSA1 <- subset (THWSA_post1986, Date > as.Date("1986-12-01"), select = -c(GEO,Date,year,month))
THWNSA1 <- subset (THWNSA_post1986, select = -c(GEO,Date,year,month))
info_columns <- subset (THWNSA_post1986, select = c(GEO,Date,year,month))

THWSF_post1986 <- cbind(info_columns, THWSA1/THWNSA1)

# get ride of spaces in province names

THWSF_post1986 = THWSF_post1986 %>%
mutate(GEO=str_replace_all(GEO,pattern=" ", replacement="."))

EMPSF_post1986 = EMPSF_post1986 %>%
mutate(GEO=str_replace_all(GEO,pattern=" ", replacement="."))

THW_post1986 = THW_post1986 %>%
mutate(GEO=str_replace_all(GEO,pattern=" ", replacement="."))

province_index = select(EMPSF_post1986, GEO)%>%unique()


```

## We want to reshape the long SF tables into  wide table 

```{r}
EMPSF_post1986 <- pivot_wider(EMPSF_post1986, names_from = "GEO", values_from =c(contains(".")))

EMPSF_post1986 <- EMPSF_post1986 %>% select(-contains("year"))
EMPSF_post1986 <- EMPSF_post1986 %>% select(-contains("month"))

#

THWSF_post1986 <- pivot_wider(THWSF_post1986, names_from = "GEO", values_from =c(contains(".")))

THWSF_post1986 <- THWSF_post1986 %>% select(-contains("year"))
THWSF_post1986 <- THWSF_post1986 %>% select(-contains("month"))

# 

THW_post1986 <- pivot_wider(THW_post1986, names_from = "GEO", values_from =c(contains(".")))

THW_post1986 <- THW_post1986 %>% select(-contains("year"))
THW_post1986 <- THW_post1986 %>% select(-contains("month"))




# Take THW_post1986 dataframe, multiply it by the ratio of two columns in the EMPSF_post1986 dataframe - (NAICS_prov/NAICS_Canada)*THW_post1986
# make a for loop over province


eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("EMPSF_ratio = EMPSF_post1986 %>% select( paste0({THW.variables},'_', province_index, '/', {THW.variables},'_', Canada)"))
     )
     )
           


eval(parse(text=
     # view the line of code generated by running with the glue command        
             paste0(EMPSF_ratio = EMPSF_post1986 %>% select( paste0(glue("{THW.variables}_Ontario"), '/', glue("{THW.variables}_Canada"))))
        )
     )
     

  
```




## Now, using the employment and hours seasonal factors, we can create a final seasonal adjustment factor for each province - note, THWSF is only at the national level

```{r}

THW_test <- subset (THW_post1986, Date > as.Date("2021-12-01"), select = c(Date, Agriculture..111.112..1100..1151.1152._Canada, Agriculture..111.112..1100..1151.1152._Ontario, Agriculture..111.112..1100..1151.1152._Quebec))

THWSF_test <- subset (THWSF_post1986, Date > as.Date("2021-12-01"), select = c(Date, Agriculture..111.112..1100..1151.1152._Canada))


EMPSF_test <- subset (EMPSF_post1986, Date > as.Date("2021-12-01"), select = c(Date, Agriculture..111.112..1100..1151.1152._Canada, Agriculture..111.112..1100..1151.1152._Ontario, Agriculture..111.112..1100..1151.1152._Quebec))

dftest <- THW_test["Agriculture..111.112..1100..1151.1152._Ontario"]*THWSF_test["Agriculture..111.112..1100..1151.1152._Canada"]*EMPSF_test[3]/EMPSF_test[2]

THW_test <- subset (THW_post1986, Date > as.Date("2021-12-01"), select = c(Date, select(c(contains("Agriculture.", "Manufacturing.")))))


select(THW_test,matches("Ontario") & matches ("Agriculture"))

clist <- c("Ontario", "Quebec")
for (i in clist) {
THW_test[i] <- select(THW_test,contains(i))
   }


select(c(-contains("Canada")))
select(c(contains("Canada")))




SF_post1986 = THWSF_post1986

```
 

### Check which variables were now left out from crafting our total
Compare these to the list in the previous chunk that was used to build **totals.hours.ours**.  Note that above, some of these _leftover_ variables were already used (eg. mining + agriculture).



## Save the modified data sets:
```{r}

THW_post1986      %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-total_hours_worked.csv"))
THWSA_post1986      %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-total_hours_worked_SA.csv"))
EMPSA_post1986      %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-employment_SA.csv"))
EMPNSA_post1986      %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-employment_NSA.csv"))
GDP_post1996       %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-GDP.csv"))
GDPAnnual_post1996 %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-GDPAnnual.csv"))
GDP_pre1997       %>% write_csv(file = paste0(location,"1.c-Data_From_Cansim-GDPretro.csv"))

save("THW_post1986", "THWSA_post1986", "EMPSA_post1986", "EMPNSA_post1986",  "GDP_post1996", "GDPAnnual_post1996", "GDP_pre1997",
     "THW.variables", "GDP.variables", "GDPSIC.variables",
file = paste0(location,"1.c-Data_From_Cansim-GDP_and_total_hours_worked.Rdata"))
```


# To do from here

- Obtain the GDP share per sector for a given year.
- Convert that into GDP share per province per sector using _Total Hours Worked_.