---
title: "1.GDP_complete"
author: "Raven Wheesk"
date: "6/17/2022"
output: html_document
---

# Data Provider

Statistics Canada hosts a large set of socio-economic [datasets](https://www150.statcan.gc.ca/n1//en/type/data?MM=1#tables). 


# Basic libraries and R version

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(dplyr)  # for data manipulation

# set up data location:
location = "C:/users/wher/OneDrive - Bank of Canada/MyDocs/Google_Climate_Dataset/"
```

```{R version}
version
```

# Loading data
```{r loaddata, echo = FALSE}
# loading from here to avoid re-downloading
load("Total_hours_and_GDP.Rdata")
```

#Variable lists
```{r andrevalues, cache = TRUE}

NAICS_higher_levels = c(
"All industries [T001]",
"Agriculture, forestry, fishing and hunting [11]",
"Mining, quarrying, and oil and gas extraction [21]",	
"Utilities [22]",
"Construction [23]",
"Manufacturing [31-33]",
"Wholesale trade [41]",
"Retail trade [44-45]",
"Transportation and warehousing [48-49]",
"Information and cultural industries [51]",
"Finance and insurance [52]",
"Real estate and rental and leasing [53]",
"Professional, scientific and technical services [54]",
"Management of companies and enterprises [55]",
"Administrative and support, waste management and remediation services [56]",
"Educational services [61]",
"Health care and social assistance [62]",
"Arts, entertainment and recreation [71]",
"Accommodation and food services [72]",
"Other services (except public administration) [81]",
"Public administration [91]")

GDP.variables = c(
"Utilities..22.",
"Construction..23.",                                  
"Manufacturing..31.33.",  
"Transportation.and.warehousing..48.49.", 
"Educational.services..61.",                      
"Health.care.and.social.assistance..62.",
"Accommodation.and.food.services..72.",
"Other.services..except.public.administration...81.",
"Public.administration..91.",
"Agriculture..forestry..fishing.and.hunting..11.mining.21",
"Wholesale.and.retail.trade..41..44.45.",
"Finance..insurance..real.estate..rental.and.leasing..52.53.",
"Professional..scientific.and.technical.services..54.", 
"Information..culture.and.recreation..51..71.")

SIC_NAICS_common = c(
"All industries",
"Agricultural and related services industries",
"Fishing and trapping industries",
"Logging and forestry industries",
"Mining (including milling), quarrying and oil well industries",
"Manufacturing industries",
"Construction industries",
"Transportation and storage industries",
"Communication industries",
"Other utility industries",
"Wholesale trade industries",
"Retail trade industries",
"Finance, insurance and real estate industries",
"Business services industries",
"Government service industries",
"Educational service industries",
"Health and social service industries",
"Accommodation, food and beverage service industries",
"Other service industries",
"Amusement and recreational service industries",
"Postal and courier service industries")

GDPsic.variables = c(
"Agriculture..forestry..fishing.and.hunting..11.mining.21",
"Manufacturing.industries",
"Construction.industries",
"Transportation.and.warehousing..48.49.",
"Information..culture.and.recreation..51..71.",
"Other.utility.industries",
"Wholesale.and.retail.trade..41..44.45.",
"Finance..insurance.and.real.estate.industries",
"Business.services.industries",
"Government.service.industries",
"Educational.service.industries",
"Health.and.social.service.industries",
"Accommodation..food.and.beverage.service.industries",
"Other.service.industries") 

GDPannual.variables = c(
"All.industries..T001.",
"Utilities..22.", 
"Construction..23.", 
"Manufacturing..31.33.",    
"Transportation.and.warehousing..48.49.", 
"Educational.services..61.", 
"Health.care.and.social.assistance..62.",
"Accommodation.and.food.services..72.", "Other.services..except.public.administration...81.", "Public.administration..91.", "Agriculture..forestry..fishing.and.hunting..11.mining.21", "Wholesale.and.retail.trade..41..44.45.", "Finance..insurance..real.estate..rental.and.leasing..52.53.", "Professional..scientific.and.technical.services..54.",  "Information..culture.and.recreation..51..71.","All.industries.ours")

```

## First we clean the monthly GDP data (NAICS)

```{r exploreGDP, cache=TRUE}
GDP_wide = 
  GDP %>% 
  filter(GEO =="Canada")%>%
  filter(Seasonal.adjustment=="Seasonally adjusted at annual rates")%>%
  filter(Prices=="Chained (2012) dollars")%>%
  filter(North.American.Industry.Classification.System..NAICS. %in%NAICS_higher_levels)%>%
  select(GEO,Date,year, month,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the NAICS names
```

## Extending Management back to 1997

```{r}
#value of management in dec 2006 x 1 - growth rate of gdp in January 2007 

#create variable for GDP growth rate:
GDP_wide = GDP_wide %>% 
   arrange(Date) %>%
mutate(GDP_growth=All.industries..T001./lag(All.industries..T001.)-1) 

#(1) Looping backward from 2006-12-01:

start     <- as.Date("06-12-01",format="%y-%m-%d")
end   <- as.Date("97-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_wide = GDP_wide %>% mutate(Management.of.companies.and.enterprises..55. = case_when( (Date < as.Date("2007-01-01")) ~ lead(Management.of.companies.and.enterprises..55.)*(1-lead(GDP_growth)),(Date >= as.Date("2007-01-01")) ~ Management.of.companies.and.enterprises..55.))
  theDate <- theDate - 1}

```

# Aggregating some NAICS categories to match SIC and LFS
```{r}
GDP_wide = GDP_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21   =  Agriculture..forestry..fishing.and.hunting..11. +Mining..quarrying..and.oil.and.gas.extraction..21.)%>%
  mutate(Wholesale.and.retail.trade..41..44.45.                     = Wholesale.trade..41.+Retail.trade..44.45.)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. = Finance.and.insurance..52.+Real.estate.and.rental.and.leasing..53.)%>%
  mutate(Information..culture.and.recreation..51..71.                = Arts..entertainment.and.recreation..71.+Information.and.cultural.industries..51.)%>%
  mutate(Professional..scientific.and.technical.services..54.       = Management.of.companies.and.enterprises..55.+Administrative.and.support..waste.management.and.remediation.services..56.+Professional..scientific.and.technical.services..54.)

#Dropping unnecessary columns
GDP_wide <- GDP_wide %>% select("GEO", "Date", "year", "month",(contains({GDP.variables})))

# Add sum column:
eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDP_wide = GDP_wide %>%  mutate(All.industries.ours = ", paste({GDP.variables},collapse='+'),")"))
           )
     )
```

## Next, we clean the monthly GDP data (SIC)
```{r}
GDPsic_wide = 
  GDPsic %>% 
  filter(Seasonal.adjustment=="Seasonally adjusted")%>%
  filter(Standard.Industrial.Classification..1980..SIC. %in% SIC_NAICS_common)%>%
  select(GEO,Date,year, month,
               Standard.Industrial.Classification..1980..SIC.,
               VALUE)%>%
  pivot_wider(names_from= Standard.Industrial.Classification..1980..SIC., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the SIC names
```


# Aggregating some SIC categories to match NAICS and LFS
```{r}
GDPsic_wide = GDPsic_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21 = Agricultural.and.related.services.industries + Fishing.and.trapping.industries + Logging.and.forestry.industries + Mining..including.milling...quarrying.and.oil.well.industries)%>%
  mutate(Wholesale.and.retail.trade..41..44.45. = Wholesale.trade.industries + Retail.trade.industries)%>%
  mutate(Transportation.and.warehousing..48.49. = Transportation.and.storage.industries + Postal.and.courier.service.industries)%>%
  mutate(Information..culture.and.recreation..51..71. = Communication.industries - Postal.and.courier.service.industries + Amusement.and.recreational.service.industries) %>%
mutate(Other.service.industries = Other.service.industries - Amusement.and.recreational.service.industries)

#Dropping unnecessary columns
GDPsic_wide <- GDPsic_wide %>% select("GEO", "Date", "year", "month",(contains({GDPsic.variables})))

# Add sum column:
eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDPsic_wide = GDPsic_wide %>%  mutate(All.industries.ours = ", paste({GDPsic.variables},collapse='+'),")"))
           )
     )
```

## Mapping industry names for the SIC codes to NAICS codes

```{r}
#Changing column names in GDPsic_wide to match column names in GDP_wide
GDPsic_wide = GDPsic_wide %>%
  mutate(Utilities..22. = Other.utility.industries)%>%
  mutate(Construction..23. = Construction.industries)%>%
  mutate(Manufacturing..31.33. = Manufacturing.industries)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. =
Finance..insurance.and.real.estate.industries)%>%
  mutate(Professional..scientific.and.technical.services..54. = Business.services.industries)%>%
  mutate(Public.administration..91. = Government.service.industries)%>%
  mutate(Educational.services..61. = Educational.service.industries)%>%
  mutate(Health.care.and.social.assistance..62. = Health.and.social.service.industries)%>%
  mutate(Accommodation.and.food.services..72. = Accommodation..food.and.beverage.service.industries)%>%
  mutate(Other.services..except.public.administration...81. = Other.service.industries)

#Dropping unnecessary columns
GDPsic_wide <- GDPsic_wide %>% select("GEO", "Date", "year", "month", (contains({GDP.variables})),"All.industries.ours")

#Changing range of GDPsic_wide:
GDPsic_wide=GDPsic_wide %>% filter(Date >= "1975-12-01" & Date <= "1997-01-01")

```

## Splicing GDPsic_wide and GDP_wide - GDP_sic uses constant prices, and GDP_wide uses 2012 chained prices. To splice them, we need to shift the level of the GDPsic_wide to that of GDP_wide. 
# We do this by creating a blank dataframe with the date range that we want (1976-1997).

```{r}
dates<-seq(from=as.Date("1975-12-01","19%y-%m-%d"),to=as.Date("1996-12-01","19%y-%m-%d"),by="month")

cols <- c('GEO', 'Date', 'year', 'month',GDP.variables, 'All.industries.ours')

GDP_blank <- data.frame(matrix(1, nrow = length(dates), ncol = 19))
names(GDP_blank) <- cols
GDP_blank$Date <- dates

GDP_wide_test <- rbind(GDP_blank, GDP_wide)
GDP_wide_test["GEO"] = "Canada"

GDP_wide_test["year"] = format(GDP_wide_test["Date"],format="%Y")
GDP_wide_test["month"] = format(GDP_wide_test["Date"],format="%m")

```

#Now we do the same to GDPsic_wide, but with a blank dataframe from 1997 onward. 
```{r}
dates<-seq(from=as.Date("1997-02-01","19%y-%m-%d"),to=as.Date("2022-03-01","20%y-%m-%d"),by="month")

cols <- c('GEO', 'Date', 'year', 'month',GDP.variables, 'All.industries.ours')

GDP_blank1 <- data.frame(matrix(1, nrow = length(dates), ncol = 19))
names(GDP_blank1) <- cols
GDP_blank1$Date <- dates

GDPsic_wide_test <- rbind(GDPsic_wide,GDP_blank1)
GDPsic_wide_test["GEO"] = "Canada"

#GDPsic_wide_test["Date"] = GDP_wide_test["Date"]

#format(GDPsic_wide_test["Date"], format="%y-%m-%d")

GDPsic_wide_test["year"] = format(GDP_wide_test["Date"],format="%Y")
GDPsic_wide_test["month"] = format(GDP_wide_test["Date"],format="%m")

```

#We use full_join to join the two tables together.
```{r}
GDPfinal_wide <- GDPsic_wide_test %>%
  full_join(GDP_wide_test, by=c("Date", "GEO", "year", "month"),all=TRUE)

```

#Now we splice the data using a backward loop starting in 1997. 
```{r}
#(1) Looping backward from 2006-12-01:
start     <- as.Date("97-01-01",format="%y-%m-%d")
end   <- as.Date("75-12-01",format="%y-%m-%d")

coltest = c(paste0(GDP.variables,".y",sep=""),"All.industries.ours.y")
coltest1 = c(paste0(GDP.variables,".x",sep=""),"All.industries.ours.x")

counter = 1

for (counter in c(1:length(coltest))){
  theDate <- start
while (theDate >= end){
  GDPfinal_wide = GDPfinal_wide %>% mutate(across(coltest[counter], ~ case_when( (Date < as.Date("97-01-01", format="%y-%m-%d")) ~ get(coltest1[counter])/lead(get(coltest1[counter]))*lead(get(coltest[counter])),(Date >= as.Date("97-01-01", format="%y-%m-%d")) ~ get(coltest[counter]))))
  theDate <- theDate - 1}
}

#Dropping unnecessary columns
GDPfinal_wide <- GDPfinal_wide %>% select("GEO", "Date", "year", "month",(c(paste0(GDP.variables,".y",sep=""),"All.industries.ours.y")))

colnames(GDPfinal_wide)<-gsub(".y","",colnames(GDPfinal_wide))

```

## We also need GDP by province by industry annual for calculating productivity later. Here we clean the table, 
```{r exploreGDP, cache=TRUE}

GDPAnnual %>%
select(Value)%>% table %>% names %>% sort

GDPAnnual_wide = 
  GDPAnnual %>% 
  filter(Value=="Chained (2012) dollars")%>%
  filter(North.American.Industry.Classification.System..NAICS. %in%NAICS_higher_levels)%>%
  select(GEO,Date,
               North.American.Industry.Classification.System..NAICS.,
               VALUE)%>%
  pivot_wider(names_from= North.American.Industry.Classification.System..NAICS., 
              values_from=VALUE) %>%  # puts each province into its own column
  rename_all(make.names) # rebuilt the NAICS names

#aggregating some NAICS
GDPAnnual_wide = GDPAnnual_wide %>%
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21   =  Agriculture..forestry..fishing.and.hunting..11. +Mining..quarrying..and.oil.and.gas.extraction..21.)%>%
  mutate(Wholesale.and.retail.trade..41..44.45.                     = Wholesale.trade..41.+Retail.trade..44.45.)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. = Finance.and.insurance..52.+Real.estate.and.rental.and.leasing..53.)%>%
  mutate(Professional..scientific.and.technical.services..54.       = Management.of.companies.and.enterprises..55.+Administrative.and.support..waste.management.and.remediation.services..56.+Professional..scientific.and.technical.services..54.)%>%
  mutate(Information..culture.and.recreation..51..71.                = Arts..entertainment.and.recreation..71.+Information.and.cultural.industries..51.)

#Obtaining a sum
eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDPAnnual_wide = GDPAnnual_wide %>%  mutate(All.industries.ours = ", paste({GDP.variables},collapse='+'),")"))
           )
     )

GDPAnnual_wide <- GDPAnnual_wide[c("GEO", "Date", GDPannual.variables)]

GDPAnnual_wide <- pivot_wider(GDPAnnual_wide, names_from = "GEO", values_from=c(GDPannual.variables))

```

## Save the modified data sets:

```{r}
GDPfinal_wide       %>% write_csv(file = paste0(location,"1.GDPfinal_wide.csv"))
GDPAnnual_wide %>% write_csv(file = paste0(location,"1.GDPAnnual_wide.csv"))

save("GDPfinal_wide", "GDPAnnual_wide", file = paste0(location,"1.GDP.Rdata"))
```
