---
title: "1.Productivity"
author: "Raven Wheesk"
date: "6/17/2022"
output: html_document
---

# Data Provider

Statistics Canada hosts a large set of socio-economic [datasets](https://www150.statcan.gc.ca/n1//en/type/data?MM=1#tables). 

# Objective: Use the monthly, provincial GDP by industry calculated in 1.GDP_complete, the seasonally adjusted total hours worked by province and industry, and the labour share of hours worked by province and industry to calculate provincial GDP for each industry. 


# Basic libraries and R version

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(dplyr)  # for data manipulation
library(data.table)
# set up data location:
location = "C:/users/wher/OneDrive - Bank of Canada/MyDocs/Google_Climate_Dataset/"
```

```{R version}
version
```

# Loading data
```{r loaddata, echo = FALSE}
# loading from here to avoid re-downloading
#load("Total_hours_and_GDP.Rdata")
load("1.LFS.Rdata")
load("1.GDP.Rdata")
```

#Variable lists:
```{r}
GDP.variables = c(
"Utilities..22.",
"Construction..23.",                                  
"Manufacturing..31.33.",  
"Transportation.and.warehousing..48.49.", 
"Educational.services..61.",                      
"Health.care.and.social.assistance..62.",
"Accommodation.and.food.services..72.",
"Other.services..except.public.administration...81.",
"Public.administration..91.",
"Agriculture..forestry..fishing.and.hunting..11.mining.21",
"Wholesale.and.retail.trade..41..44.45.",
"Finance..insurance..real.estate..rental.and.leasing..52.53.",
"Professional..scientific.and.technical.services..54.", 
"Information..culture.and.recreation..51..71.")

#This includes our total, "All.industries.ours"
GDP.variables1 = c(
"Utilities..22.", 
"Construction..23.", 
"Manufacturing..31.33.",    
"Transportation.and.warehousing..48.49.", 
"Educational.services..61.", 
"Health.care.and.social.assistance..62.",
"Accommodation.and.food.services..72.", "Other.services..except.public.administration...81.", "Public.administration..91.", "Agriculture..forestry..fishing.and.hunting..11.mining.21", "Wholesale.and.retail.trade..41..44.45.", "Finance..insurance..real.estate..rental.and.leasing..52.53.", "Professional..scientific.and.technical.services..54.",  "Information..culture.and.recreation..51..71.",
"All.industries.ours")
```

## To calculate annual productivity (GDPAnnual_wide/THWSA_wide), we need to convert THWSA_wide to annual. Here we also clean up the tables so that they have identical dimensions for dividing. 
```{r}
#Note since we're calculating annual productivity, we restrict the range to the most recent full year.
GDPfinal_wide1=GDPfinal_wide %>% filter(Date >= "1987-01-01" & Date <= "2021-12-01")
THWSA_wide1=THWSA_wide %>% filter(Date >= "1986-12-01" & Date <= "2021-12-01")
info_columns <- subset (GDPfinal_wide1, select = c(year))
Date_column <- subset (GDPfinal_wide1, select = c(Date, year, month))

THWSA_wide1 = cbind(info_columns,THWSA_wide1)

#Generate our THWSA annual dataframe + tidy up column names
THWSAAnnual_wide <- THWSA_wide1 %>%
  group_by(year) %>% 
  dplyr::summarize(across(everything(), list(mean))) %>% 
  as.data.frame()

THWSAAnnual_wide =  THWSAAnnual_wide %>% rename_with(~str_remove(., '_1'))
THWSAAnnual_wide=THWSAAnnual_wide %>% filter(Date >= "1997-01-01" & Date <= "2021-12-01")
#Drop territories
GDPAnnual_wide1 <- GDPAnnual_wide %>% select("Date", (contains({GDP.variables1})), -contains(c("_Nunavut", "_Yukon", "_Northwest Territories")))
THWSAAnnual_wide1 <- THWSAAnnual_wide %>% select(-contains("year"))

# obtain a column of NATIONAL totals within each NAICS group for each year. We do this by creating a sum. Since we sum across rows but still want our Canada column, we make Canada = 0 to avoid double counting (note, this can probably be done a different way).
GDPAnnual_wide1 =
eval(parse(text=
glue(paste0(
'GDPAnnual_wide1 %>%
group_by(Date)%>% # within a year
mutate(',paste(paste0(GDP.variables1,"_Canada","=0,"),collapse=","),')'))
))%>%
ungroup()

col1 = c(paste0(GDP.variables1,"_Canada",sep=""))
col2 = c(GDP.variables1)
counter = 1
for (counter in c(1:length(col1))){
  GDPAnnual_wide1 = GDPAnnual_wide1 %>% mutate(across(col1[counter], ~ rowSums(select(GDPAnnual_wide1,contains({col2[counter]})))))

}
```

#Calculating productivity:
```{r}
columnlist = c(colnames(THWSAAnnual_wide1))

GDPAnnual_wide1 <- GDPAnnual_wide1[,columnlist]

info_columns_annual <- subset (GDPAnnual_wide2, select = c(Date))
GDPAnnual_wide1 = subset(GDPAnnual_wide1, select = -c(Date) )
THWSAAnnual_wide1 = subset(THWSAAnnual_wide1, select = -c(Date) )

#Take the annual GDP and THW dataframes and create a productivity dataframe
productivity = GDPAnnual_wide1/THWSAAnnual_wide1*1000

productivity = cbind(info_columns_annual,productivity)

```

#Now we want to extend our productivity dataframe back to 1987. For convenience right now, we assume productivity remains constant between 1987 and 1997. Note that we could use a different approach here.
```{r}
dates<-seq(from=as.Date("1987-07-01","19%y-%m-%d"),to=as.Date("1996-07-01","19%y-%m-%d"),by="year")
cols <- c(columnlist)

productivity_blank <- data.frame(matrix(1, nrow = length(dates), ncol = 166))
names(productivity_blank) <- cols
productivity_blank$Date <- dates

productivity1 = rbind(productivity_blank,productivity)
columnlist1 = columnlist[-1]

start     <- as.Date("97-07-01",format="%y-%m-%d")
end   <- as.Date("87-07-01",format="%y-%m-%d")

for (counter in c(2:length(columnlist1))){
  theDate <- start
while (theDate >= end){
  productivity1 = productivity1 %>% mutate(across(columnlist[counter], ~ case_when( (Date < as.Date("97-07-01", format="%y-%m-%d")) ~ lead(get(columnlist[counter])),(Date >= as.Date("97-07-01", format="%y-%m-%d")) ~ get(columnlist[counter]))))
  theDate <- theDate - 1}
}

```

# Now that we have annual productivity, we need to convert it to monthly for calculating monthly GDP. 
```{r}
#Add year and month variables
DT <- productivity1
setDT(DT)
DT[, Date := as.IDate(Date)]
DT[, year := year(Date)]
setkey(DT, year)
productivity2 <- DT[CJ(year, month = 1:12)]
print(productivity2)

class(productivity2) <- class(as.data.frame(productivity2))

productivity2 <- productivity2 %>% filter(Date <= "2021-12-01") %>% select(-contains(c("Date","year","month")))

THWSA_share1 <- THWSA_share %>% filter(Date <= "2021-12-01") %>% select(-contains(c("Date")))

#Adjusting productivity by the labour share - check this
productivity_share <- productivity2*THWSA_share1

GDP.variables.list <- c(GDP.variables1)

col1 = c(paste0(GDP.variables.list,"_Canada",sep=""))
col2 = c(GDP.variables.list)
counter = 1
for (counter in c(1:length(col1))){
  productivity_share = productivity_share %>% mutate(across(col1[counter], ~ rowSums(select(productivity_share,contains({col2[counter]})))-select(productivity_share,contains({col1[counter]}))))

}

unnest_dataframes <- function(x) {

        y <- do.call(data.frame, x)

        if("data.frame" %in% sapply(y, class)) unnest_dataframes(y)

        y

    }

    productivity_share <- unnest_dataframes(productivity_share)
#Cleaning up names
names(productivity_share) <- gsub("Newfoundland.and.Labrador", "Newfoundland and Labrador", names(productivity_share))
names(productivity_share) <- gsub("Prince.Edward.Island", "Prince Edward Island", names(productivity_share))
names(productivity_share) <- gsub("Nova.Scotia", "Nova Scotia", names(productivity_share))
names(productivity_share) <- gsub("New.Brunswick", "New Brunswick", names(productivity_share))
names(productivity_share) <- gsub("British.Columbia", "British Columbia", names(productivity_share))
```

# Productivity varies across industries and provinces. We need to calculate the relative productivity between provinces and Canada for a given industry to adjust our estimates. 
```{r}    
productivity_CAN <- productivity_share %>% select(contains("Canada"))
##
#GDP.variables.list <- c(GDP.variables1)
col1 = c(paste0(GDP.variables.list,"_Canada",sep=""))
col2 = c(GDP.variables.list)
counter = 1
for (counter in c(1:length(col1))){
  productivity_prov = productivity_prov %>% mutate(across(col1[counter], ~ select(productivity_CAN,contains({col1[counter]}))))

}
#Cleaning up names
productivity_prov <- unnest_dataframes(productivity_prov)
names(productivity_prov) <- gsub("Newfoundland.and.Labrador", "Newfoundland and Labrador", names(productivity_prov))
names(productivity_prov) <- gsub("Prince.Edward.Island", "Prince Edward Island", names(productivity_prov))
names(productivity_prov) <- gsub("Nova.Scotia", "Nova Scotia", names(productivity_prov))
names(productivity_prov) <- gsub("New.Brunswick", "New Brunswick", names(productivity_prov))
names(productivity_prov) <- gsub("British.Columbia", "British Columbia", names(productivity_prov))

productivity_CAN <- data.frame(do.call(cbind,rep(productivity_CAN, each=11)), check.names = F)
productivity_ratio <- productivity_prov/productivity_CAN
productivity_ratio <- cbind(Date_column, productivity_ratio)
```
### Finally, we calculate our GDP by NAICS by month with: TWHSA_wide * THWSA_share * productivity_ratio
```{r}
productivity_ratio1 <- productivity_ratio %>% select(-contains(c("Date", "year", "month")))
GDPfinal_wide1 <- GDPfinal_wide1 %>% select(-contains(c("GEO","Date", "year", "month")))
GDPfinal_wide1 <- data.frame(do.call(cbind,rep(GDPfinal_wide1, each=11)), check.names = F)
THWSA_share1 <- THWSA_share %>% filter(Date <= "2021-12-01") %>% select(-contains("Date"))

finalGDP_bys_byp <- GDPfinal_wide1*(THWSA_share1)*productivity_ratio1

finalGDP_bys_byp <- cbind(Date_column, finalGDP_bys_byp)
colnames(finalGDP_bys_byp) <- c(colnames(productivity_ratio))
```

# For Dave, we pivot_long the table to have columns for GEO, and each NAICS
```{r}
finalGDP_bys_byp_long <- pivot_longer(finalGDP_bys_byp, -c(Date,year, month), names_to = c("NAICS", "GEO"), names_sep = "_", values_to = "values") %>% 
  #dplyr::select(-question) %>% 
  dplyr::arrange(year, month)

finalGDP_bys_byp_long = finalGDP_bys_byp_long %>% pivot_wider(names_from= NAICS, values_from=values) 
```

#Saving final output
```{r}
finalGDP_bys_byp_long       %>% write_csv(file = paste0(location,"finalGDP.csv"))
finalGDP_bys_byp            %>% write_csv(file=paste0(location, "finalGDPwide.csv"))
```
