---
title: "2.b_GDP_THW_mapping"
author: "Sandrine Njoo"
date: "03/11/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This file maps the names of the GDP cariables to those in the Total Hours Worked data set.

# Limitations:
The THW and GDP variables are crafted in tile **1.c**, but the code is copied over.  Potentially this could be saved from **1.c** and then reloaded here if the list is expanded.



## Creating a mapping from the LFS to Census data industries

Loading the important libraries and noting the R version.



```{r}

version

```



```{r library, warning = FALSE, message = FALSE}
#using R:

library(tidyverse)  # data manipulation and piping
library(data.table) # renaming columns in bulk using 'setnames'
library(plyr)       # recoding factor levels using 'mapvalues' 
library(lubridate)

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"

```

# load data to match

```{r, message = FALSE, warning = FALSE}

THW_post1986 = read_csv(file = paste0(location,"1.c-Data_From_Cansim-total_hours_worked.csv"))
THW_post1986 %>% glimpse
GDP_post1996 = read_csv(file = paste0(location, "1.c-Data_From_Cansim-GDP.csv"))
GDP_post1996 %>% glimpse
GDP_pre1997= read_csv(file = paste0(location, "1.c-Data_From_Cansim-GDPretro.csv"))
GDP_pre1997 %>% glimpse


```


## Creating a mapping from GDP to THW data
First consider the THW and GDP variables crafted in tile **1.c**.  The code is copied over.  Potentially this could be saved from **1.c** and then reloaded here if the list is expanded.

```{r}


THW_ind_names = c(
  "Utilities..22.",
  "Construction..23.",
  "Manufacturing..31.33.",
  "Wholesale.and.retail.trade..41..44.45.",
  "Transportation.and.warehousing..48.49.",
  "Finance..insurance..real.estate..rental.and.leasing..52.53.",
  "Professional..scientific.and.technical.services..54.", 
  "Business..building.and.other.support.services..55.56.",
  "Educational.services..61.",
  "Health.care.and.social.assistance..62.",
  "Information..culture.and.recreation..51..71.",               
  "Accommodation.and.food.services..72.",
  "Other.services..except.public.administration...81.",             
  "Public.administration..91.",
  "Agriculture..forestry..fishing.and.hunting..11.mining.21")


GDP_ind_names = c("Utilities..22.",
      "Construction..23.",                                      
      "Manufacturing..31.33.",  
      "Transportation.and.warehousing..48.49.", 
      "Professional..scientific.and.technical.services..54.",
      "Educational.services..61.",                      
      "Health.care.and.social.assistance..62.",
      "Accommodation.and.food.services..72.",
      "Other.services..except.public.administration...81.",
      "Public.administration..91.",
      "Agriculture..forestry..fishing.and.hunting..11.mining.21",
      "Wholesale.and.retail.trade..41..44.45.",
      "Finance..insurance..real.estate..rental.and.leasing..52.53.", 
      "Information..culture.and.recreation..51..71.")


dict_thw_gdp <- data.frame(GDP_ind_names=NA, THW_ind_names=NA)[numeric(0), ]
not_matched <- data.frame(not_found=NA)

for (val in 1:length(GDP_ind_names)){
  if(!identical(agrep(GDP_ind_names[[val]], THW_ind_names), integer(0))){
    out_1 = GDP_ind_names[[val]]
    out_2 = THW_ind_names[agrep(GDP_ind_names[[val]], THW_ind_names)]
    if(length(out_2)<2){
      dict_thw_gdp[val,1] <- out_1
      dict_thw_gdp[val,2] <- out_2
    }
    else{
      if(adist(GDP_ind_names[[val]],out_2[1])>adist(GDP_ind_names[[val]],out_2[2])){
        dict_thw_gdp[val,1] <- out_1
        dict_thw_gdp[val,2]<-out_2[2]
      }
      else{
        dict_thw_gdp[val,2]<-out_2[1]
      }
      
    }
    
  }
  else if(identical(agrep(GDP_ind_names[[val]], THW_ind_names), integer(0))){
    out = GDP_ind_names[[val]]
    not_matched = rbind(not_matched, out)
  }
  
}

dict_thw_gdp <- dict_thw_gdp %>% drop_na()

dict_thw_gdp[nrow(dict_thw_gdp) + 1,] = c("All.industries.ours","totals.hours.ours")
```

## Creating a mapping from GDP (1961-1997) to THW data

```{r}


THW_ind_names = c(
  "Utilities..22.",
  "Construction..23.",
  "Manufacturing..31.33.",
  "Wholesale.and.retail.trade..41..44.45.",
  "Transportation.and.warehousing..48.49.",
  "Finance..insurance..real.estate..rental.and.leasing..52.53.",
  "Professional..scientific.and.technical.services..54.", 
  "Business..building.and.other.support.services..55.56.",
  "Educational.services..61.",
  "Health.care.and.social.assistance..62.",
  "Information..culture.and.recreation..51..71.",               
  "Accommodation.and.food.services..72.",
  "Other.services..except.public.administration...81.",             
  "Public.administration..91.",
  "Agriculture..forestry..fishing.and.hunting..11.mining.21")

#For GDPsic_ind_names, changing names manually may be necessary for easier matching


GDP_pre1997 = GDP_pre1997 %>%
  mutate(Utilities..22. = Other.utility.industries)%>%
  mutate(Construction..23. = Construction.industries)%>%
  mutate(Manufacturing..31.33. = Manufacturing.industries)%>%
  mutate(Finance..insurance..real.estate..rental.and.leasing..52.53. =
Finance..insurance.and.real.estate.industries)%>%
  mutate(Professional..scientific.and.technical.services..54. = Business.services.industries)%>%
  mutate(Public.administration..91. = Government.service.industries)%>%
  mutate(Educational.services..61. = Educational.service.industries)%>%
  mutate(Health.care.and.social.assistance..62. = Health.and.social.service.industries)%>%
  mutate(Accommodation.and.food.services..72. = Accommodation..food.and.beverage.service.industries)%>%
  mutate(Other.services..except.public.administration...81. = Other.service.industries)

  #
  
GDPsic_ind_names = c(
  "Manufacturing..31.33.",
  "Construction..23.",
  "Utilities..22.",
  "Finance..insurance..real.estate..rental.and.leasing..52.53.",
  "Professional..scientific.and.technical.services..54.",
  "Public.administration..91.",
  "Educational.services..61.",
  "Health.care.and.social.assistance..62.",
  "Accommodation.and.food.services..72.",
  "Other.services..except.public.administration...81.",
  "Agriculture..forestry..fishing.and.hunting..11.mining.21",
  "Wholesale.and.retail.trade..41..44.45.",
  "Transportation.and.warehousing..48.49.",
  "Information..culture.and.recreation..51..71.")


dict_thw_gdpsic <- data.frame(GDPsic_ind_names=NA, THW_ind_names=NA)[numeric(0), ]
not_matched <- data.frame(not_found=NA)

for (val in 1:length(GDPsic_ind_names)){
  if(!identical(agrep(GDPsic_ind_names[[val]], THW_ind_names), integer(0))){
    out_1 = GDPsic_ind_names[[val]]
    out_2 = THW_ind_names[agrep(GDPsic_ind_names[[val]], THW_ind_names)]
    if(length(out_2)<2){
      dict_thw_gdpsic[val,1] <- out_1
      dict_thw_gdpsic[val,2] <- out_2
    }
    else{
      if(adist(GDPsic_ind_names[[val]],out_2[1])>adist(GDPsic_ind_names[[val]],out_2[2])){
        dict_thw_gdpsic[val,1] <- out_1
        dict_thw_gdpsic[val,2]<-out_2[2]
      }
      else{
        dict_thw_gdpsic[val,2]<-out_2[1]
      }
      
    }
    
  }
  else if(identical(agrep(GDPsic_ind_names[[val]], THW_ind_names), integer(0))){
    out = GDPsic_ind_names[[val]]
    not_matched = rbind(not_matched, out)
  }
  
}

dict_thw_gdpsic <- dict_thw_gdpsic %>% drop_na()

dict_thw_gdpsic[nrow(dict_thw_gdpsic) + 1,] = c("All.industries.ours","totals.hours.ours")

#note - gdpsic not fully mapping to thw - used to mutate to change them


```

# Joining the two dictionaries

```{r}

dict_thw_gdp_gdpsic <- inner_join(dict_thw_gdp, dict_thw_gdpsic)

```

# Examine dictionary mapping

```{r}

dict_thw_gdp_gdpsic

```



## Checking for failures:

```{r}
not_matched
```

## Save the output

```{r}
dict_thw_gdp_gdpsic %>% write_csv(file= paste0(location, "2.b_Industry_mapping_THW_Census_GDP_THW_GDPSIC.csv"))
```

# Use dictionary mapping to create GDP dataframes with identical columns - this section can be relocated

```{r}

naics_gdp_vars <- full_dict_new %>% pull(GDP_ind_names) %>% unique() 
sic_gdp_vars <- full_dict_new %>% pull(GDPsic_ind_names) %>% unique() 

new_GDP_post1996 <- GDP_post1996 %>% select(GEO, Date, all_of(naics_gdp_vars))
new_GDP_pre1997 <- GDP_pre1997 %>% select(GEO, Date, all_of(sic_gdp_vars))



```

#now have identical table structure for GDP tables for 1961-
#Next, create data frame that splices pre1997 GDP to post1996 GDP. Options include: (1) Create new table from GDP_post1996, extend range backward to 1976. Create new table from GDP_pre1997 that is the industry GDP growth rates. Use formula from excel table to splice each column by growth rate table. 

```{r}


#create variable for GDP growth rate for each industry:

GDP_1976 <- new_GDP_pre1997 %>% filter(Date >= "1975-12-01" & Date <= "1997-01-01")


GDP_1976 = GDP_1976 %>% 
   arrange(Date) %>%
mutate(Utilities_g=Utilities..22./lead(Utilities..22.),
Construction_g=Construction..23./lead(Construction..23.), 
Manufacturing_g=Manufacturing..31.33./lead(Manufacturing..31.33.), 
Transportation_g=Transportation.and.warehousing..48.49./lead(Transportation.and.warehousing..48.49.),
Professional_g=Professional..scientific.and.technical.services..54./lead(Professional..scientific.and.technical.services..54.),
Educational_g=Educational.services..61./lead(Educational.services..61.),
Health.care_g=Health.care.and.social.assistance..62./lead(Health.care.and.social.assistance..62.),
Accommodation_g=Accommodation.and.food.services..72./lead(Accommodation.and.food.services..72.),
Other.services_g=Other.services..except.public.administration...81./lead(Other.services..except.public.administration...81.),
Public.administration_g=Public.administration..91./lead(Public.administration..91.),
Agriculture_g=Agriculture..forestry..fishing.and.hunting..11.mining.21/lead(Agriculture..forestry..fishing.and.hunting..11.mining.21),
Trade_g=Wholesale.and.retail.trade..41..44.45./lead(Wholesale.and.retail.trade..41..44.45.),
Finance_g=Finance..insurance..real.estate..rental.and.leasing..52.53./lead(Finance..insurance..real.estate..rental.and.leasing..52.53.),
Information_g=Information..culture.and.recreation..51..71./lead(Information..culture.and.recreation..51..71.))

#note: there must be an easier way to write this with the lapply function


#to bind this table to the post_1996 table, need matching columns, including growth rate columns:


match_GDP_post1996 = new_GDP_post1996 %>% 
   arrange(Date) %>%
mutate(Utilities_g=Utilities..22./lead(Utilities..22.),
Construction_g=Construction..23./lead(Construction..23.), 
Manufacturing_g=Manufacturing..31.33./lead(Manufacturing..31.33.), 
Transportation_g=Transportation.and.warehousing..48.49./lead(Transportation.and.warehousing..48.49.),
Professional_g=Professional..scientific.and.technical.services..54./lead(Professional..scientific.and.technical.services..54.),
Educational_g=Educational.services..61./lead(Educational.services..61.),
Health.care_g=Health.care.and.social.assistance..62./lead(Health.care.and.social.assistance..62.),
Accommodation_g=Accommodation.and.food.services..72./lead(Accommodation.and.food.services..72.),
Other.services_g=Other.services..except.public.administration...81./lead(Other.services..except.public.administration...81.),
Public.administration_g=Public.administration..91./lead(Public.administration..91.),
Agriculture_g=Agriculture..forestry..fishing.and.hunting..11.mining.21/lead(Agriculture..forestry..fishing.and.hunting..11.mining.21),
Trade_g=Wholesale.and.retail.trade..41..44.45./lead(Wholesale.and.retail.trade..41..44.45.),
Finance_g=Finance..insurance..real.estate..rental.and.leasing..52.53./lead(Finance..insurance..real.estate..rental.and.leasing..52.53.),
Information_g=Information..culture.and.recreation..51..71./lead(Information..culture.and.recreation..51..71.))



GDP_1976 <- GDP_1976 %>% filter(Date >= "1976-01-01" & Date < "1997-01-01")
GDP_total <- rbind(GDP_1976, match_GDP_post1996)



#create full GDP table by splicing pre_1997 according to growth rate

#now test on this table to match to excel sheet - repeat for all 14 variables. Then, create a final GDP_full table with only the 14 variables

eval(parse(text=
     # view the line of code generated by running with the glue command       
             glue(paste0('GDP_wide %>%arrange(Date)%>% mutate(', paste(paste0(GDP.variables,"_","g","=",GDP.variables, "/", lead(GDP.variables),collapse=","),')')
                        ))))

# use glue to calculate productivity
 eval(parse(text= glue("productivity_annual_avg = productivity_annual_avg %>% 
mutate({paste('prod', {naics_thw_vars}, sep = '_')} = (as.numeric({paste0({naics_thw_vars},'.y')}) /(as.numeric({paste0({naics_thw_vars},'.x')})))*1000)")))
 
 GDP_wide_test <- GDP_wide %>%
 
 eval(parse(text= glue('GDP_wide = GDP_wide %>%arrange(Date)%>% mutate(, paste({paste("g", {GDP.variables}, sep= "_")} = (as.numeric({paste0({GDP.variables})})/(as.numeric({paste0(lead{GDP.variables})}')
                        ))

#Utilities:

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Utilities..22. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Utilities..22.)*(Utilities_g),(Date >= as.Date("1997-01-01")) ~ Utilities..22.))
  theDate <- theDate - 1}

##Now for the rest of the columns - will do the brute force way for now, can clean up later

#Construction:

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Construction..23. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Construction..23.)*(Construction_g),(Date >= as.Date("1997-01-01")) ~ Construction..23.))
  theDate <- theDate - 1}

# Manufacturing:

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Manufacturing..31.33. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Manufacturing..31.33.)*(Manufacturing_g),(Date >= as.Date("1997-01-01")) ~ Manufacturing..31.33.))
  theDate <- theDate - 1}

# Transportation and warehousing:

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start


while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Transportation.and.warehousing..48.49. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Transportation.and.warehousing..48.49.)*(Transportation_g),(Date >= as.Date("1997-01-01")) ~ Transportation.and.warehousing..48.49.))
  theDate <- theDate - 1}


# Professional, scientific and technical services:

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Professional..scientific.and.technical.services..54. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Professional..scientific.and.technical.services..54.)*(Professional_g),(Date >= as.Date("1997-01-01")) ~ Professional..scientific.and.technical.services..54.))
  theDate <- theDate - 1}

#Educational services

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start


while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Educational.services..61. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Educational.services..61.)*(Educational_g),(Date >= as.Date("1997-01-01")) ~ Educational.services..61.))
  theDate <- theDate - 1}

# Health care and social assistance

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start


while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Health.care.and.social.assistance..62. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Health.care.and.social.assistance..62.)*(Health.care_g),(Date >= as.Date("1997-01-01")) ~ Health.care.and.social.assistance..62.))
  theDate <- theDate - 1}

# Accommodation and food services

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start


while (theDate >= end){
  GDP_total = GDP_total %>% mutate(Accommodation.and.food.services..72. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Accommodation.and.food.services..72.)*(Accommodation_g),(Date >= as.Date("1997-01-01")) ~ Accommodation.and.food.services..72.))
  theDate <- theDate - 1}

# Other services

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Other.services..except.public.administration...81. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Other.services..except.public.administration...81.)*(Other.services_g),(Date >= as.Date("1997-01-01")) ~ Other.services..except.public.administration...81.))
  theDate <- theDate - 1} 

#Public administration

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Public.administration..91. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Public.administration..91.)*(Public.administration_g),(Date >= as.Date("1997-01-01")) ~ Public.administration..91.))
  theDate <- theDate - 1} 

#Agriculture, forestry, fishing, hunting and mining

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Agriculture..forestry..fishing.and.hunting..11.mining.21 = case_when( (Date < as.Date("1997-01-01")) ~ lead(Agriculture..forestry..fishing.and.hunting..11.mining.21)*(Agriculture_g),(Date >= as.Date("1997-01-01")) ~ Agriculture..forestry..fishing.and.hunting..11.mining.21))
  theDate <- theDate - 1} 

# Wholesale and retail trade

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Wholesale.and.retail.trade..41..44.45. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Wholesale.and.retail.trade..41..44.45.)*(Trade_g),(Date >= as.Date("1997-01-01")) ~ Wholesale.and.retail.trade..41..44.45.))
  theDate <- theDate - 1} 

#Finance and insurance

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Finance..insurance..real.estate..rental.and.leasing..52.53. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Finance..insurance..real.estate..rental.and.leasing..52.53.)*(Finance_g),(Date >= as.Date("1997-01-01")) ~ Finance..insurance..real.estate..rental.and.leasing..52.53.))
  theDate <- theDate - 1} 

# Information, culture and recreation

start     <- as.Date("96-12-01",format="%y-%m-%d")
end   <- as.Date("76-01-01",format="%y-%m-%d")

theDate <- start

while (theDate >= end){
  GDP_total = GDP_total %>% mutate (Information..culture.and.recreation..51..71. = case_when( (Date < as.Date("1997-01-01")) ~ lead(Information..culture.and.recreation..51..71.)*(Information_g),(Date >= as.Date("1997-01-01")) ~ Information..culture.and.recreation..51..71.))
  theDate <- theDate - 1} 




```

# All columns are now matched. A final table can be made dropping the growth rate variables and 
```{r}
eval(parse(text=
     # view the line of code generated by running with the glue command        
             glue(paste0("GDP_full = GDP_total %>%  mutate(All.industries.ours = ", paste({GDP_ind_names},collapse='+'),")"))
           )
     )

GDP_full <- GDP_full %>% select(-contains("_g"))
      
```




## Save the modified data set:
```{r}

GDP_full      %>% write_csv(file = paste0(location,"2.d-Data_From_Cansim-final_gdp.csv"))

```