---
title: "2.c_Industry_mapping_LFS_Census_GDP_THW"
author: "Sandrine Njoo"
date: "03/11/2021"
output: html_document
---

This file defines the mapping from the LFS NAICS coded variables to those from the census GDP and the Total Hours Worked datasets.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(stringr)
library(RColorBrewer) # for colours

# set up data location:
location = "C:/Users/wher/OneDrive - Bank of Canada/MyDocs/git"
```


```{r}

version
```





## Creating a mapping between all 4 datasets



We already have a mapping between LFS and Census and between THW and GDP from files **2.1** and **2.b**
```{r, message = FALSE, warning = FALSE}

dict_new     = read_csv(file = paste0(location, "2.a-LFT_to_census_dictionary.csv"))

dict_thw_gdp = read_csv(file = paste0(location, "2.b_Industry_mapping_THW_Census_GDP_THW.csv"))

```


First, we'll extract the NAICS code (2 digit integer) to then combine our two dict_newionaries together. 



```{r}
# find the naics code numbers:
dict_new = dict_new %>% mutate(naics =  str_replace_all(census,pattern = "[^\\d]+", replacement = "_"))
# remove introduced leading or tailing punctuation:
dict_new = dict_new %>% mutate(naics =  str_replace_all(naics,pattern = "(^_)|(_$)", replacement = ""))

# find the naics code numbers:
dict_thw_gdp = dict_thw_gdp %>% mutate(naics =  str_replace_all(THW_ind_names,pattern = "[^\\d]+", replacement = "_"))
# remove introduced leading or tailing punctuation:
dict_thw_gdp = dict_thw_gdp %>% mutate(naics =  str_replace_all(naics,pattern = "(^_)|(_$)", replacement = ""))

```

# Merge the two dictionaries based on the NAICS codes

Note that some mappings are many to one since some NAICS industries have been merged together.  

```{r}
full_dict_new <- dict_thw_gdp%>% inner_join(dict_new, by = "naics")

```


# Colour mapping
At this time it is convenient to tie in a column of colours to each variable to keep them consistent across all plots later on.
```{r}
#define the mapping
names2use = unique(full_dict_new$GDP_ind_names)
# include some extras in case more values are added later
color2use = unique(c(brewer.pal(n=7,name = "Set1"),
                     brewer.pal(n=8, name = "Set2"),
                     brewer.pal(n=8, name = "Set3")))[1:length(names2use)]


# set up one colour per merged NAICS code
full_dict_new <- full_dict_new %>% 
 mutate(colours2use = plyr::mapvalues(GDP_ind_names,   from = names2use,      to = color2use))


full_dict_new %>% glimpse
```

Created the full dict_newionary for the 4 datasets
```{r}
full_dict_new %>% write_csv( file = paste0(location, '2.c_full_dict_mapping_THW_GDP_LFS_Census.csv'))
# previously saved as read.csv( 'full_dict_new.csv')
```