---
title: "3.b Dividing monthly GDP into provincial share by NAICS code"
author: "Sandrine Njoo"
date: "04/11/2021"
output: html_document
---

Take national GDP by NAICS industry and split it into provincial amounts based on total hours worked within a province.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Set up the environment
```{R}
version
```


Load libraries.
```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(stringr)

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```

Read relevant CSVs
```{r, message = FALSE}
full_dict_new <- read_csv(paste0(location, "2.c_full_dict_mapping_THW_GDP_LFS_Census.csv"))
THW_post2007  <- read_csv(paste0(location, "1.c-Data_From_Cansim-total_hours_worked.csv"))
GDP_post2007  <- read_csv(paste0(location, "1.c-Data_From_Cansim-GDP.csv"))
```

Remove the columns we don't want (for example Agriculture, forestry etc since we already have agriculutre and mining together). We get a subset of the dataframes
```{r}
naics_gdp_vars <- full_dict_new %>% filter(naics !=  "N/A" )%>% pull(GDP_ind_names) %>% unique() 
naics_thw_vars <- full_dict_new %>% filter(naics !=  "N/A" )%>% pull(THW_ind_names) %>% unique() 

new_THW_post2007 <- THW_post2007 %>% select(GEO, Date, all_of(naics_gdp_vars))
new_gdp_post2007 <- GDP_post2007 %>% select(GEO, Date, all_of(naics_thw_vars))

```



Deal with missing values in the  Agriculutre and utilities columns by subtracting the Canadian total from the sum of the values within other provinces for a given date.  In all cases there is at most one missing value per row.

```{r}
#new_THW_post2007_prov[which(is.na(new_THW_post2007_prov), arr.ind=TRUE)]

new_THW_post2007 <- new_THW_post2007 %>% 
  group_by(Date) %>% 
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21 = 
           ifelse (is.na(Agriculture..forestry..fishing.and.hunting..11.mining.21), abs(-2*max(Agriculture..forestry..fishing.and.hunting..11.mining.21,na.rm=TRUE)+ sum(Agriculture..forestry..fishing.and.hunting..11.mining.21,na.rm=TRUE)),
                   Agriculture..forestry..fishing.and.hunting..11.mining.21 )) %>%
  mutate(Utilities..22. = ifelse (is.na(Utilities..22.), 
                                  abs(-2*max(Utilities..22.,na.rm=TRUE)+sum(Utilities..22.,na.rm=TRUE)),
                                  Utilities..22. ))%>%
  ungroup()

# The sum includes the Canadian total.  Remove Canada from the sum (it's the maximum).  
# Then take [Canada - sum of others] as the point estimate.

```

We add a new column with the total hours across all sectors for a given location.  There should no longer be any NA values.

```{r}
new_THW_post2007 = new_THW_post2007 %>% 
                 mutate(totals.hours.ours = rowSums(new_THW_post2007%>% select(-GEO, -Date)))
```

We remove the rows with Canada and we get the ratio of hours worked in each province per sector

```{r}
scale2 <- function(x){x/sum(x)}


new_THW_post2007 <-   new_THW_post2007 %>% 
                          filter(GEO !="Canada") %>%
                          group_by(Date) %>% 
                          mutate(across(all_of(naics_thw_vars),scale2))%>%# all our NAICS encoded variables end with numbers
                          ungroup()

```

We multiply each ratio by the corresponding gdp per month per sector, by iterating through the dates and through the sectors.

First merge the national and provincial datasets based on Date.  This copies rows of national information into each province.  Then take products of the provincial share and the national amount.

```{r, echo = TRUE, eval = TRUE}

# merge the two data sets to make multiplying easier at a fixed date
# This also ensures the all the dates align:
Merged_gdp_thw = new_gdp_post2007%>% select(-GEO) %>% inner_join(new_THW_post2007, by = "Date")
# define a placeholder:
new_THW_split = Merged_gdp_thw%>% select(GEO, Date, totals.hours.ours) # just keep the location, date, and total hours

# use glue to match the NAICS variables
# populate the NAICS variable columns:
eval(parse(text =
             glue(paste0("new_THW_split = new_THW_split %>% mutate({naics_thw_vars} = Merged_gdp_thw${naics_thw_vars}",".x*Merged_gdp_thw${naics_thw_vars}",".y)"))
          )
     )



```


## Take a look




```{r}
new_THW_split%>% glimpse
```
# Save output
 
```{r}
new_THW_split %>% write_csv(file = paste0(location, '3.b_GDP_contribution_per_NAICS_province_post2007.csv'))
# formerly known as: "THW_GDP_post2007.csv"

```