---
title: '3.c - GDP Into Provincial Share'
author: "Dave Campbell"
date: '2021-11-08'
summary: 'GDP share'
lastmod: '2021-11-01T16:34:12-04:00'
---




# Goals

Using GDP at the {national, monthly, NAICS sector} determine **GDP** share at {provincial, monthly, NAICS sector}.  We want to be able to assess provinces with larger / smaller share of the NAICS sector.




# Basic libraries and R version

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```

```{R version}
version
```

  
  
## Loading the GDP tables

```{r loading, message = FALSE, warning = FALSE}

# # GDP monthly, national:
GDP_post2007       = read_csv(file = paste0(location, "1.c-Data_From_Cansim-GDP.csv"))
# # GDP annual, provincial:
GDPAnnual_post2007 = read_csv(file = paste0(location, "1.c-Data_From_Cansim-GDPAnnual.csv"))
full_dict_new <- read_csv(file = paste0(location, '2.c_full_dict_mapping_THW_GDP_LFS_Census.csv'))

naics_gdp_vars <- full_dict_new %>% filter(naics !=  "N/A" )%>% pull(GDP_ind_names) %>% unique() 
naics_thw_vars <- full_dict_new %>% filter(naics !=  "N/A" )%>% pull(THW_ind_names) %>% unique() 

```



# Time stamps

Check the resolution of the two datasets:

```{r}

GDPAnnual_post2007 %>% select(Date)%>% table
GDP_post2007 %>% select(Date)%>% table %>% head(14)

```

Years in the annual data set are time stamped as July 1 and applied for the full calendar year.



# Obtain annual fraction of GDP for \{sector, province\}

Using annual data obtain the fraction of GDP within a sector that is allocated to each province



```{r, warnings = FALSE}

GDP_annual_national_by_NAICS = GDP_post2007 %>% 
  select(year,all_of(naics_gdp_vars)) %>%
  group_by(year)%>%
  summarize(across(everything(),sum))%>% ungroup()


# want to convert GDP within a NAICS into GDP fraction per province

GDP_Totals_annual_provincial_by_NAICS = 
  eval(parse(text=
     # obtain a column of NATIONAL totals within each NAICS group for each year:
   glue(paste0(
    'GDPAnnual_post2007 %>% 
    group_by(Date)%>% # within a year
    mutate(',paste(paste0("Total_All_Provinces.",naics_gdp_vars,"=sum(",naics_gdp_vars,")"),collapse=","),')'))
))%>%
  ungroup()
```


### confirm this acted as intended:

These should be equal.
```{r}
GDP_Totals_annual_provincial_by_NAICS %>% 
  filter(Date == "2007-07-01")%>% 
  select(Agriculture..forestry..fishing.and.hunting..11.mining.21)%>% sum
GDP_Totals_annual_provincial_by_NAICS %>% 
  filter(Date == "2007-07-01")%>%
  select(Total_All_Provinces.Agriculture..forestry..fishing.and.hunting..11.mining.21)%>% unique()
```


## Obtain the provincial fraction of {national, annual, GDP per sector}  


```{r}
# complete the fraction
GDP_fraction_annual_provincial_by_NAIC =
  eval(parse(text=
  # obtain a column of PROVINCIAL FRACTIONS within each NAICS group for each year:
glue(paste0(
  ' GDP_Totals_annual_provincial_by_NAICS %>% mutate(',paste(paste0("Fraction_Annual_Provinces.",naics_gdp_vars,"=",naics_gdp_vars,"/","Total_All_Provinces.",naics_gdp_vars),collapse=","),')'))
))
```



### confirm this acted as intended:

These should be equal and within a date it should sum to 1.  Note that dates here are actually _year_ since they are all July 1st for a given year.
```{R}
GDP_Totals_annual_provincial_by_NAICS     %>% filter(Date == "2007-07-01")%>% select(                          Agriculture..forestry..fishing.and.hunting..11.mining.21)/
   GDP_Totals_annual_provincial_by_NAICS  %>% filter(Date == "2007-07-01")%>% select(                          Agriculture..forestry..fishing.and.hunting..11.mining.21)%>% sum
GDP_fraction_annual_provincial_by_NAIC    %>% filter(Date == "2007-07-01")%>% select(Fraction_Annual_Provinces.Agriculture..forestry..fishing.and.hunting..11.mining.21, GEO)
GDP_fraction_annual_provincial_by_NAIC    %>% filter(Date == "2007-07-01")%>% select(Fraction_Annual_Provinces.Agriculture..forestry..fishing.and.hunting..11.mining.21)%>%sum #sums to 1

```



# Obtain the monthly GDP allocation for \{sector, province\}
Using annual fraction of GDP per \{province, sector\} assume that GDP per sector maintains a constant fractional allocation per \{province, month\}.

Calculate the monthly GDP per \{sector, province\} by taking the annual fraction of NAICS sector within a province and applying that fraction to monthly national GDP:



```{r}
# craft the building blocks:
GDP_FRACTIONS_annual_provincial_by_NAIC = GDP_fraction_annual_provincial_by_NAIC%>% 
   separate(Date, into = c("year","month","day")) %>% 
  mutate(year = as.numeric(year))%>%
   select(GEO, year, paste0("Fraction_Annual_Provinces.",naics_gdp_vars))


GDP_monthly_national_by_NAICS = GDP_post2007 %>% select(Date, year, month, all_of(naics_gdp_vars))



# join these together then multiply the fraction allocated to a province to the monthly GDP total
 GDP_total_monthly_by_province_sector = eval(parse(text=
     # obtain a column of NATIONAL totals within each NAICS group for each year:
   glue(paste0(
     'GDP_FRACTIONS_annual_provincial_by_NAIC %>% 
     inner_join(GDP_monthly_national_by_NAICS) %>%
     group_by(year)%>%
     mutate(',paste(paste0("GDP_allocation_for_province.",naics_gdp_vars,"=Fraction_Annual_Provinces.",naics_gdp_vars,"*",naics_gdp_vars),collapse=","),')'))
))%>%
  ungroup()%>%
   select(Date, GEO,starts_with('GDP_allocation_for_province')) # cleanup

 
 

```


# Try some plots

```{r plots}


GDP_total_monthly_by_province_sector %>% 
  pivot_longer(cols = -c(Date, GEO), names_to= "sector", values_to="GDP")%>%
  ggplot()+
geom_line(aes(x=Date, y = GDP, colour = sector),lwd=2,alpha = .5)+
  facet_wrap(~GEO)+ylim(0,110000)+
  ggtitle("Monthly GDP per Province per Sector")



GDP_FRACTIONS_annual_provincial_by_NAIC %>% 
  select(year, GEO,starts_with('Fraction_Annual_Provinces.'))%>%
  pivot_longer(cols = -c(year, GEO), names_to= "sector", values_to="GDP")%>%
  ggplot()+
  geom_line(aes(x=year, y = GDP, colour = GEO, group = GEO),lwd=2,alpha = .5)+
  facet_wrap(~sector)+
  ggtitle("Fraction Within Sector Allocated to Each Province")

```



# saving output
```{r saving}
# the GDP total for a province in a month in a sector:
GDP_total_monthly_by_province_sector  %>% write_csv(file = paste0(location, "3.c-GDP_total_monthly_by_province_sector.csv"))
GDP_Totals_annual_provincial_by_NAICS %>% write_csv(file = paste0(location, "3.c-GDP_Totals_annual_provincial_by_NAICS.csv"))


# some interim results:
save("GDP_FRACTIONS_annual_provincial_by_NAIC",
    "GDP_annual_national_by_NAICS",
    "GDP_monthly_national_by_NAICS",
    "GDP_total_monthly_by_province_sector", file = paste0(location, "3.c-GDP_total_monthly_by_province_sector.Rdata"))

```



