---
title: "3.d Converting THW and Annual GDP into Productivity"
author: "Sandrine Njoo & Dave Campbell"
date: "09/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Goal
Use the total hours worked within a sector and province to determine the amount of effort.  Obtain the productivity per sector , province, month by taking the GDP and slicing it based on effort.


# Set up the work environment

```{r}
version
```

Libraries
```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(cansim) #for pulling Stat Can socioeconomic data
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(stringr)

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```

Load the data that will be used
```{r, message = FALSE, warning = FALSE}

full_dict_new <- read_csv(paste0(location, "2.c_full_dict_mapping_THW_GDP_LFS_Census.csv"))
THW_post2007 <- read_csv(paste0(location, "1.c-Data_From_Cansim-total_hours_worked.csv"))
GDP_Totals_annual_provincial_by_NAICS <- read_csv(paste0(location,'3.c-GDP_Totals_annual_provincial_by_NAICS.csv'))
```




Extract the columns we want from the GDP and THW dataframes.  Make sure that _year_ is included
```{r}
# This code is also used in 3.c to get the provincial share 

naics_thw_vars = unique(full_dict_new$THW_ind_names)
naics_gdp_vars = unique(full_dict_new$GDP_ind_names)

# Just the raw GDP per {sector, year, province}
new_gdp_total_post2007 = GDP_Totals_annual_provincial_by_NAICS %>%
  select(GEO, Date, all_of(naics_gdp_vars))%>% 
         mutate(year = year(Date))

#Total hours worked per {month, year, sector, province}
new_THW_post2007 = THW_post2007 %>%
    select(GEO, Date, year,all_of(naics_thw_vars))


```

Replace NA values.  Deal with missing values in the  Agriculutre and utilities columns by subtracting the Canadian total from the sum of the values within other provinces for a given date.  In all cases there is at most one missing value per row.


```{r}
new_THW_post2007 <- new_THW_post2007 %>% 
  group_by(Date) %>% 
  mutate(Agriculture..forestry..fishing.and.hunting..11.mining.21 = 
           ifelse (is.na(Agriculture..forestry..fishing.and.hunting..11.mining.21),          
                   abs(-2*max(Agriculture..forestry..fishing.and.hunting..11.mining.21,na.rm=TRUE)+sum(Agriculture..forestry..fishing.and.hunting..11.mining.21, na.rm=TRUE)),
                   Agriculture..forestry..fishing.and.hunting..11.mining.21 )) %>%
  mutate(Utilities..22. = 
           ifelse (is.na(Utilities..22.), 
                   abs(-2*max(Utilities..22.,na.rm=TRUE)+sum(Utilities..22.,na.rm=TRUE)),
                   Utilities..22. ))%>%
  ungroup()

```


# Obtain Provincial Monthly Productivity by sector


Sum total hours across all industries for a {province, month, year}.
```{r}
new_THW_post2007 = new_THW_post2007 %>% 
  group_by(GEO,Date,year)%>%
  mutate(totals.hours.ours = sum(c_across(naics_thw_vars))) %>% 
  ungroup

```


Obtain the average hours worked for the year per province per sector.
Data is annualized so each month is an estimate for the year.  The average month is then the average estimate for the year.

```{r}



scale2 <- function(x, na.rm = FALSE) mean(x)
new_averageTHW_post2007 = new_THW_post2007 %>% 
  filter(GEO !="Canada") %>%
  group_by(year, GEO) %>% 
  mutate(across(all_of(naics_thw_vars),scale2))%>%# all our NAICS encoded variables end with numbers
  ungroup()
                          

```

### Define productivity
Productivity is the yearly GDP for each province and sector divided by the annual estimated hours worked.



```{r}
productivity_annual_avg <- new_averageTHW_post2007 %>%
  inner_join(new_gdp_total_post2007, by=c("year", "GEO"))

# use glue to calculate productivity
 eval(parse(text= glue("productivity_annual_avg = productivity_annual_avg %>% 
mutate({paste('prod', {naics_thw_vars}, sep = '_')} = (as.numeric({paste0({naics_thw_vars},'.y')}) /(as.numeric({paste0({naics_thw_vars},'.x')})))*1000)")))



```


Remove unwanted columns
```{r}
productivity_annual_avg = productivity_annual_avg %>% 
  mutate(Date = Date.x)%>% 
  select(c(GEO, Date, year), starts_with('prod'), ends_with('ours'))
```

Note that this produces a constant average productivity value repeated across each month.  Use this later to split it into monthly portions of effort.
```{r}
productivity_annual_avg %>% glimpse
```

Get the proportion of hours worked for each province in each sector compared to the national number.
```{r}
# same as code from:
# 3.b_GDP_THW_census_grouping.Rmd
scale3 <- function(x){x/sum(x)}
new_THW_post2007_ratio  = new_THW_post2007  %>% 
                          filter(GEO !="Canada")
new_THW_post2007_ratio  = new_THW_post2007_ratio %>% 
                          group_by(Date) %>% 
                          mutate(across(naics_thw_vars,scale3))%>%
                          ungroup()

```

Multiply the proportion of hours worked within each province by the productivity for the appropriate province and sector 

```{r}
prod_by_prov <-  new_THW_post2007_ratio%>% inner_join(productivity_annual_avg, by = c("Date", "GEO"))

for (var_name in naics_thw_vars){
 eval(parse(text= glue("prod_by_prov = prod_by_prov %>% 
mutate({paste('gdp_prod', {var_name}, sep = '_')} = (as.numeric({var_name}) *(as.numeric({paste0('prod_',{var_name})}))))")))
}

```

Remove unwanted column for the production 
```{r}
prod_by_prov <- prod_by_prov %>% select(c(Date, GEO), starts_with('gdp_prod_'))
```

```{r}
prod_by_prov %>% glimpse
```

# Obtain National Monthly Productivity by sector
Obtain national production by sector
```{r}
prod_national <- prod_by_prov %>% 
  group_by(Date) %>% 
  dplyr::summarise(across(where(is.numeric), sum))%>% 
  mutate(GEO = "Canada")

```

```{r}
prod_national %>%
  glimpse
```


# Save output

```{r}

productivity_annual_avg   %>% write_csv( paste0(location, "3.d_productivity_per_province.csv"))
prod_national             %>% write_csv( paste0(location, "3.d_productivity_national.csv"))

```


```{r, eval = FALSE, echo = FALSE}

#
# productivity_annual_avg_old = read_csv( "3.d-productivity_per_province_deprecated.csv") 
# test = productivity_annual_avg %>% inner_join(productivity_annual_avg_old, by = c("GEO","Date"))
# # 
#  lp=0
# lp=lp+1
# var2use = paste0('prod_',naics_thw_vars[lp])#
# # var2use = 'totals.hours.ours'# this is the only different column
# tmp = eval(parse(text = glue(paste0("cbind(test${var2use}",".x",",test${var2use}",".x-","test${var2use}",".y)"))))
# summary(tmp)
# 
# 
# 
# 
# 
# prod_national_old           = read_csv('3.d-productivity_national_deprecated.csv')
# test = prod_national %>% inner_join(prod_national_old, by = c("GEO","Date"))
# lp=0
# lp=lp+1
# var2use = paste0('gdp_prod_',naics_thw_vars[lp])#
# 
# tmp = eval(parse(text = glue(paste0("cbind(test${var2use}",".x",",test${var2use}",".x-","test${var2use}",".y)"))))
# summary(tmp)

```




