---
title: "3.e THW and GDP adjusted"
author: "Sandrine Njoo & Dave Campbell"
date: "15/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal
Use productivity and split it into provincial share

# Set up the work environment

```{r}
version
```

Libraries

```{R, message = FALSE, warning = FALSE}
library(tidyverse) # for its nice flow.
library(lubridate) # for dealing with dates
library(GGally) # for pairs plot
library(ggpubr) # for putting a regression equation onto a ggplot figure
library(glue)   # used for dynamic variable construction
library(stringr)

# set up data location:
location = "R:/res/wher/Google_Climate_Dataset/"
```
## 

# Data loading

Load and examine the data that will be used.  Note the differences in time range and geographies.  Information from Territories is not available.

```{r, message = FALSE, warning = FALSE}
# unique values of GDP per {NAICS, month, province}
THW_GDP_post2007 = read_csv(paste0(location, '3.b_GDP_contribution_per_NAICS_province_post2007.csv'))
THW_GDP_post2007 %>% pull(Date)%>% range
THW_GDP_post2007 %>% pull(GEO)%>% unique
THW_GDP_post2007 %>% glimpse


# average productivity per {NAICS, year, province} but the same annual value is repeated once per month to make merging and maths easier below
productivity_annual_avg     = read_csv(paste0(location, "3.d_productivity_per_province.csv"))
# productivity_annual_avg   = read_csv(paste0(location, "3.d-productivity_per_province_deprecated.csv"))
productivity_annual_avg %>% pull(Date)%>% range
productivity_annual_avg %>% pull(GEO)%>% unique
productivity_annual_avg %>% glimpse


# unique productivity per {NAICS, month, national} 
prod_national   = read_csv(paste0(location, "3.d_productivity_national.csv"))
# prod_national = read_csv(paste0(location, "3.d-productivity_national_deprecated.csv"))
prod_national %>% pull(Date)%>% range
prod_national %>% pull(GEO)%>% unique
prod_national %>% glimpse


# for mapping NAICS variables to colours
full_dict_new <- read_csv(paste0(location, "2.c_full_dict_mapping_THW_GDP_LFS_Census.csv"))
colmat = full_dict_new%>% select(GDP_ind_names,colours2use,census)%>% unique()

```

Load a helper dictionary with variable names
```{r, message = FALSE, warning = FALSE}
full_dict_new <- read_csv(paste0(location, "2.c_full_dict_mapping_THW_GDP_LFS_Census.csv"))
naics_gdp_vars = unique(full_dict_new$GDP_ind_names)

```





# Calculate the adjusted GDP

Obtain the fraction of productivity of a province compared to the national level.  The monthly fraction of productivity is the provincial value divided by the national value.


```{r}

# merge the two data sets to make multiplying easier at a fixed date
# This ensures the all the dates align:
Merged_gdp = prod_national %>%  inner_join(productivity_annual_avg, by = "Date")


# use glue to match the NAICS variables
# populate the NAICS variable columns:
eval(parse(text =
             glue(paste0("Merged_gdp = Merged_gdp %>% mutate(gdp_ratio_{naics_gdp_vars} = prod_{naics_gdp_vars}","/gdp_prod_{naics_gdp_vars})"))
          )
     )


# define a place to keep the useful info:
ratios = Merged_gdp%>% select(GEO.y, Date, starts_with("gdp_ratio_"))%>% # just keep the location and date
  dplyr::rename("GEO" = "GEO.y")


```
```{r}
ratios %>% glimpse
```

# Split productivity into the fraction of gdp within each province
Multiply monthly fraction of productivity per province by the values in THW_GDP_post2007. Do this via a similar merging strategy to ensure GEO and Dates match.
Generate the code automatically using **glue** so that all variables are appropriately constructed.  To reconstruct the generated lines of code used run the 2 lines for the **glue** command.

```{r}
productivity_adjusted_GDP_post2007 <- merge(THW_GDP_post2007, ratios, by = c('Date',"GEO"))

eval(parse(text= 
glue("productivity_adjusted_GDP_post2007 = productivity_adjusted_GDP_post2007 %>% 
mutate({paste0('adjusted_', {naics_gdp_vars})} = (as.numeric({naics_gdp_vars}) * as.numeric({paste0('gdp_ratio_', {naics_gdp_vars})})))")
))

productivity_adjusted_GDP_post2007 <- productivity_adjusted_GDP_post2007%>% select(c('Date', 'GEO'), starts_with('adjusted_'))

productivity_adjusted_GDP_post2007%>% glimpse
```

# Save the output 
Also save a reshaped version of the output that will make plotting faster in the Shiny Apps.

```{r, eval = TRUE}

productivity_adjusted_GDP_post2007 %>% write_csv(paste0(location, "3.e_productivity_adjusted_GDP_post2007.csv"))

productivity_adjusted_GDP_post2007 %>%
  pivot_longer(starts_with("adjusted"), names_to="NAICS", values_to="productivity")%>%
  mutate(colourval = plyr::mapvalues(NAICS, from = paste0("adjusted_",colmat$GDP_ind_names), to = colmat$colours2use))%>% 
  write_csv(paste0(location, "3.e_productivity_adjusted_GDP_post2007_tall.csv"))


```



