---
title: "4.b- Productivity per NAICS within Province"
author: "Sandrine Njoo & Dave Campbell"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal
Obtain the productivity adjusted GDP per NAICS per census subdivision for a given time point.  Using the 2016 census, subdivide the monthly productivity (from **3e**) into shares based on residence within a Census Subdivision or Dissemination Area.


# Set up the work environment

```{r}
version
```

load libraries
```{r library, warning = FALSE, message = FALSE}
#using R:
library(tidyverse)  # data manipulation and piping
library(data.table) # renaming columns in bulk using 'setnames'
library(plyr)       # recoding factor levels using 'mapvalues'  
library(sf)         # to handle census subdivision polygons
library(stringr)    # look for spcific provinces and csv files 
library(glue)       # dynamic variable names
library(ggpubr)     # plotting layouts
library(lubridate)  # working with dates

# set up data location:
location = "R:/mfa/data_sci_div/Google Climate Project/"
```



## Data loading

Load and examine the data that will be used.  Note the differences in time range and geographies.  Information from Territories is not available.


```{r, message = FALSE, warning = FALSE, eval = TRUE}
# Provincial encodings:
province_codings = read_csv(file = paste0(location, '1.a_province_codings.csv'))


# dictionary for converting NAICS code names across data types:
full_dict_new <- read_csv( file = paste0(location, "2.c_full_dict_mapping_THW_GDP_LFS_Census.csv")) %>% select(GDP_ind_names,census,THW_ind_names,colours2use)%>% unique()
# variable names:
naics_gdp_vars    <- full_dict_new %>% pull(GDP_ind_names) 
naics_census_vars <- full_dict_new %>% pull(census)        
naics_thw_vars    <- full_dict_new %>% pull(THW_ind_names) 


# Fraction of the provincial totals of a NAICS sector within a CSD and geometries:
census_fraction_of_NAICS_in_csd_production = 
                readRDS(paste0(location, "4.a_census_fraction_of_subdivision_data/4.a_census_fraction_of_NAIC_in_csd_Canada.rds"))

# productivity per province
productivity_adjusted_GDP_post2007 = read_csv(paste0(location, "3.e_productivity_adjusted_GDP_post2007.csv"))


```


# Subdivide the province into CSD based on the census


The resulting data frames are very large; all census subdivisions are represented in each month.  These could be filtered and saved as individual files depending on where it is being used.

```{r , eval = FALSE}
# Consider all dates.  Alternatively uncomment:
# month = "06"
# year = "2015"
# and include filter commands such as: 
# filter(Date ==paste(year,month,"01",sep="-")) 


# Merge the productivity per {naics, month, provice} with the census fraction of province in CSD.
census_and_prod =  census_fraction_of_NAICS_in_csd_production %>%
  # filter(Date ==paste(year,month,"01",sep="-")) %>% # if filtering by date
  inner_join(productivity_adjusted_GDP_post2007, by = c("provincename" = "GEO"))

# census_and_prod is structured now with:
# variables starting with "adjusted_"   are the GDP per province, one value per province
# variables starting with "X"           are the census csd fraction of that NAICS industry per province
# variables starting with "production_in_csd" are the production per NAICS split into census csd allocation


# provincial share of production multiplied by the fraction of province in the CSD
eval(parse(text=
             glue(paste0("census_and_prod = census_and_prod %>%", 
                         "group_by(provincename, Date) %>%",
                         "mutate( production_in_csd_",'{full_dict_new$census}'," = ",
                         "adjusted_",'{full_dict_new$GDP_ind_names}',"*frac_in_csd_{full_dict_new$census}) %>%",
                         "ungroup()"))
))





# extract just the columns for productivity within {NAICS, csd, year, month} 
productivity_per_CSD = census_and_prod %>% select(Date,provincename,starts_with("production_in_csd"),Region.Name, Population, geometry)

## Dominantn NAICS based on productivity per month/year.
max_index = max.col(st_set_geometry(productivity_per_CSD%>% select(starts_with("production_in_csd_")), NULL))

productivity_per_CSD = productivity_per_CSD %>%
  mutate(Dominant_NAICS = full_dict_new$census[max_index])%>%
  mutate(colourval = plyr::mapvalues(Dominant_NAICS, 
                                     from = full_dict_new$census, 
                                     to   = full_dict_new$colours2use))
productivity_per_CSD   %>% write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_csd.rds"))

# save provincial versions of this file so that it is easier to load:
for(prov in productivity_per_CSD%>% pull(provincename) %>% unique){
  productivity_per_CSD   %>% 
    filter(provincename == prov)%>% 
    write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_csd_in",prov,".rds"))
}


# subset a test case:
year = "2020"
month = "09"
productivity_per_CSD   %>% 
      filter(Date == paste(year,month,"01",sep="-"))%>% 
  write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_csd-",year,"-",month,".rds"))




# take the productivity per province and add back in the national value
productivity_adjusted_GDP_post2007 = rbind(
                                           productivity_adjusted_GDP_post2007,
                                           productivity_adjusted_GDP_post2007 %>% 
                                               group_by(Date) %>% 
                                               dplyr::summarise(across(starts_with("adjusted"), sum)) %>% 
                                               ungroup() %>% 
                                               mutate (GEO = "Canada")
)

```


# Total productivity including national:

```{r, eval = FALSE} 
productivity_adjusted_GDP_post2007 %>% write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_per_province_updated3.ewithCanada.rds"))


```









# Repeat for Dissemination Areas



Again, the resulting data frames are very large; all census subdivisions are represented in each month.  These could be filtered and saved as individual files depending on where it is being used.

```{r, eval = FALSE}
# Consider all dates.  Alternatively uncomment the month and date info from above:
# month = "06"
# year = "2015"
# and include filter commands such as: 
# filter(Date ==paste(year,month,"01",sep="-")) 


# Fraction of the provincial totals of a NAICS sector within a DA and geometries:
census_fraction_of_NAICS_in_DA_production = 
                readRDS(paste0(location, "4.a_census_fraction_of_subdivision_data/4.a_census_fraction_of_NAIC_in_DA_Canada.rds"))

# productivity per province
productivity_adjusted_GDP_post2007 = read_csv(paste0(location, "3.e_productivity_adjusted_GDP_post2007.csv"))






# Merge the productivity per {naics, month, provice} with the census fraction of province in DA.
census_and_prod =  census_fraction_of_NAICS_in_DA_production %>%
  # filter(Date ==paste(year,month,"01",sep="-")) %>% # if filtering by date
  inner_join(productivity_adjusted_GDP_post2007, by = c("provincename" = "GEO"))

# census_and_prod is structured now with:
# variables starting with "adjusted_"   are the GDP per province, one value per province
# variables starting with "X"           are the census DA fraction of that NAICS industry per province
# variables starting with "production_in_DA" are the production per NAICS split into census DA allocation


# provincial share of production multiplied by the fraction of province in the DA
eval(parse(text=
             glue(paste0("census_and_prod = census_and_prod %>%", 
                         "group_by(provincename, Date) %>%",
                         "mutate( production_in_DA_",'{full_dict_new$census}'," = ",
                         "adjusted_",'{full_dict_new$GDP_ind_names}',"*frac_in_DA_{full_dict_new$census}) %>%",
                         "ungroup()"))
))





# extract just the columns for productivity within {NAICS, DA, year, month} 
productivity_per_DA = census_and_prod %>% select(Date,provincename,starts_with("production_in_DA"),Region.Name, Population, geometry)

## Dominantn NAICS based on productivity per month/year.
max_index = max.col(st_set_geometry(productivity_per_DA%>% select(starts_with("production_in_DA_")), NULL))

productivity_per_DA = productivity_per_DA %>%
  mutate(Dominant_NAICS = full_dict_new$census[max_index])%>%
  mutate(colourval = plyr::mapvalues(Dominant_NAICS, 
                                     from = full_dict_new$census, 
                                     to   = full_dict_new$colours2use))
productivity_per_DA   %>% write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_DA.rds"))

# save provincial versions of this file so that it is easier to load:
for(prov in productivity_per_DA%>% pull(provincename) %>% unique){
  productivity_per_DA   %>% 
    filter(provincename == prov)%>% 
    write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_DA_in",prov,".rds"))
}


# subset a test case:
year = "2020"
month = "09"
productivity_per_DA   %>% 
      filter(Date == paste(year,month,"01",sep="-"))%>% 
  write_rds(paste0(location, "4.b.ShinyApp.productivity/4.b_production_in_DA-",year,"-",month,".rds"))



```



## Next
Consider splitting the **productivity_per_CSD** and **productivity_per_DA** into pieces for year and month.

